---
title: "EP-SuRE - mESC datasets for TAD deconstruction design(s) + post-processing"
author: "FC"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
prefix.lab <- 'mesc_datasets_dload'
knitr::opts_chunk$set(echo       = TRUE,
                      message    = FALSE,
                      warning    = FALSE,
                      cache      = TRUE,
                      cache.path = paste0('cache/', prefix.lab, '/'),
                      fig.path   = paste0('figures/', prefix.lab, '/'),
                      dev        = 'png', #set to pdf for LaTeX presentation figures
                      fig.align  = 'center',
                      fig.width  = 5,
                      fig.height = 5)
```

```{r libs and fun, include = FALSE}
library(datteRo)

library(tidyverse)
library(magrittr)
library(reshape2)
library(here)
library(GEOquery)
library(UpSetR)

library(GenomicFeatures)
library(GenomicInteractions)
library(rtracklayer)
library(DESeq2)


getMetadata <- function(gse) {
  meta <- GSMList(gse)[[1]]@header
  tib  <- tibble(geo.accession = meta$geo_accession,
                 sample.id = meta$title,
                 instrument = meta$instrument_model,
                 sra.url = meta$relation[grep('SRA: ', meta$relation)])

  tib %<>%
    mutate(sra.experiment = sapply(strsplit(sra.url, 'term='), '[', 2)) %>%
    select(geo.accession, sra.experiment, everything())
  return(tib)
}

# function to read soft format metadata and return two mined metadata file for integration with Smk pipelines
extractMeta <- function(soft.fn, fq.dir, tib.lookup, raw.format = 'fastq.gz', meta.out.fn, meta.out.concat.fn) {
  
  # load soft format metadata file
  gse <- getGEO(filename = soft.fn, GSElimits = NULL, GSEMatrix = TRUE, AnnotGPL = FALSE, getGPL = FALSE)
  
  # extract accession, title & tech info
  tib.meta <- getMetadata(gse)
  
  # append sra run
  tib.meta <- tib.meta %>%
    left_join(tib.lookup, by = 'geo.accession') %>%
    select(-sra.experiment.y) %>%
    select(geo.accession, sra.run, sra.experiment = sra.experiment.x, everything())
  
  # concatenate runs for same sample
  # handles both SE and PE runs:
  ## create tibble with dumped fq filenames
  ## splitting at punctuation results in file ending with mate suffix (_1 or _2) for PE runs (mate = 1 for SE runs)
  tib.sra.run.prefix <- tibble(sra.run.fn = list.files(fq.dir))
  print(tib.sra.run.prefix)
  tib.sra.run.prefix <- tib.sra.run.prefix %>%
    mutate(sra.run = sapply(strsplit(sra.run.fn, '[[:punct:]]'), '[', 1),
           prefix = sapply(strsplit(sra.run.fn, '\\.'), '[', 1),
           mate = sapply(strsplit(prefix, '_'), '[', 2)) %>%
    replace_na(list(mate = "1"))

  # add dumped fq filenames and mate information to metadata
  tib.meta.concat <- left_join(tib.meta, tib.sra.run.prefix, by = 'sra.run')

  # generate input metadata for concatenating runs
  tib.meta.concat <- tib.meta.concat %>%
    group_by(geo.accession, mate) %>%
    mutate(sample.id = paste0(geo.accession, '_', sample.id, '_', mate),
           sra.run.files = paste0(sra.run.fn, collapse = ' ')) %>%
    ungroup %>%
    select(sample.id, sra.run.files) %>%
    distinct()
  
  # write mined metadata to files
  write_tsv(tib.meta, path = meta.out.fn)
  write_tsv(tib.meta.concat, path = meta.out.concat.fn, col_names = FALSE)
}

# read in collapsed seqmonk output and convert to GenomicInteractions object
getGI <- function(fn) {

  # read collapsed seqmonk output
  df <- read.table(fn)

  # anchors as GR --> UCSC chromosome names
  anchor.1 <- GRanges(df$V1, IRanges(df$V2, df$V3))
  seqlevelsStyle(anchor.1) <- 'UCSC'
  anchor.2 <- GRanges(df$V7, IRanges(df$V8, df$V9))
  seqlevelsStyle(anchor.2) <- 'UCSC'

  # create GI object
  gi.pchic <- GenomicInteractions(anchor1       = anchor.1,
                                  anchor2       = anchor.2,
                                  bait.name     = df$V4,
                                  otherEnd.name = df$V10,
                                  counts        = df$V5,
                                  score         = df$V6)

  # remove trans interactions
  gi.pchic <- gi.pchic[!is.trans(gi.pchic)]

  return(gi.pchic)
}

# read in collapsed seqmonk output and convert to GenomicInteractions object
get_gint_joshi <- function(df, score_col_name) {

  # anchors as GR --> UCSC chromosome names
  anchor_1 <- GRanges(df[['baitChr']], IRanges(df[['baitStart']], df[['baitEnd']]))
  anchor_2 <- GRanges(df[['oeChr']], IRanges(df[['oeStart']], df[['oeEnd']]))
  score    <- df[[score_col_name]]

  # select significant interactions (score >= 5)
  hits <- ( score >= 5 )
  
  # create GI object
  gi_pchic <- GenomicInteractions(anchor1       = anchor_1[hits],
                                  anchor2       = anchor_2[hits],
                                  score         = score[hits])
  
  return(gi_pchic)
}

# Function to read in, resize and reduce DHSs
# note: homer peak files have metadata in header of size 1 +
#   grep '#' /home/f.comoglio/Pub/GSE72164/Dnase-seq/Results/Peaks/peaks_GSM1856456_DNaseI_2i.bed | wc -l

getDHS <- function(dhs.fn, header.size = 36) {
  dhs <- ChIPseeker::readPeakFile(dhs.fn, skip = header.size)
  
  # extend to 150 nt
  dhs <- resize(dhs, 150, fix = 'center')
  
  # merge if closer than 1 nucleosome
  dhs <- reduce(dhs, min.gapwidth = 150)
     
  return(dhs)
}

# Function to combine DHS replicates
# Given set of DHS A and B, returns intersection(A, B) U (A/B) U (B/A)
combineRepl <- function(dhs.1, dhs.2) {
  int.set <- intersect(dhs.1, dhs.2)
  sdiff.1 <- subsetByOverlaps(dhs.1, dhs.2, invert = TRUE)
  sdiff.2 <- subsetByOverlaps(dhs.2, dhs.1, invert = TRUE)
  
  out <- c(int.set, sdiff.1, sdiff.2)
  
  # resize peaks < 150nt
  ext.idx <- which(width(out) < 150)
  out[ext.idx] <- resize(out[ext.idx], 150, fix = 'center')
  
  # assign score prop to evidence
  out$score <- 0.5 * countOverlaps(out, dhs.1) + countOverlaps(out, dhs.2)
  
  return(out)
}
```

---

## Aims

We plan to deconstruct selected TADs with EP-SuRE and design systematic TAD deconstruction approaches. We will use mouse ESCs (mESCs) as a developmental model system, comparing naive (pre-implantation, E4.5) and primed mESCs (post-implantation: E5.5 and E6.5 (EpiSCs)). Here, we detail all downloaded raw and processed data used for the design, and describe metadata mining for integration with our snakemake analysis workflows. In addition, we detail the post-processing of the results to generate input data sets for the DAT (Deconstruct A TAD) algorithm.

## Data source

1. All raw data from Joshi et al. (2015) Cell Stem Cell [GSE72164](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE72164). These include:

    * DNaseI-seq
    * ChIP-seq for H3K4me1/me3, H3K27ac and (H3K27me3, currently not processed)
    * RNA-seq
    * DHS CHi-C

for [E14-Tg2a](https://www.lgcstandards-atcc.org/Products/All/CRL-1821.aspx?geo_country=nl#generalinformation) mESC cultured in 2i as well as serum-LIF. Note that some experiments (part. CHi-C) have also been performed 1 and 3 days after shifting cells from serum to 2i.

2. All raw data from Factor et al. (2014) Cell Stem Cell (GEO superseries [GSE57409](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE57409). These include

    * DNaseI-seq in mEpiSC9 (129SvEv×ICR, Tesar lab) (GSE57406)
    * ChIP-seq for H3K4me1/me3, H3K27ac and H3K27me3 in mEpiSC9 (GSE57407)
    * RNA-seq (GSE57403, currently not processed) for serum-LIF E14-Tg2a mESCs and mEpiSC9
  
3. Raw DNaseI-seq profiles for serum E14-Tg2a mESCs, from John Stam's lab (released in 2012 and 2015 as part of ENCODE) [GSE37074::GSM1014154](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSM1014154)

4. Significant Promoter CHi-C interactions in serum E14-Tg2a mESCs and embryo-derived 129S2 (Brons et al. 2007) mEpiSC from Novo et al. (2018) Cell Reports [GSE103053](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE103053)

5. TAD coordinates from high-resolution Hi-C in serum E14-Tg2a mESCs from Bonev et al. (2017) Cell ([Table S2](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5651218/bin/mmc2.xlsx)) from Supporting Information, mm10)

6. All raw data from Sherwood et al. (2014) Nat Biotech [GSE53776](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE53776), which provides a DNase-seq time course profiling of serum E14-Tg2a mESCs differentiation to prepancreatic endoderm and intestinal endoderm. The profile include mesendoderm (day 3) and mesoderm and endoderm (day 5).

7. All ChIP-seq raw data from Kagey et al. (2010) Nature [GSE22562](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE22562) in [V6.5 (C57BL/6-129)](https://www.novusbio.com/products/v65-mouse-embryonic-stem-cells_nbp1-41162) mESC cultured in serum.

8. All ChIP-seq raw data from Whyte et al. (2013) Cell [GSE44286](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE44286) in V6.5 mESC cultured in serum.

9. All raw data from Galonska et al. (2015) Cell Stem Cell [GSE56312](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE56312) in [KH2 (C57BL/6 x 129/Sv)](https://web.expasy.org/cellosaurus/CVCL_C317) cultured in serum and in 2i.

10. All raw data from Chen et al. (2008) Cell [GSE11431](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE11431) in E14-Tg2a mESCs cultured in serum.

11. All raw data from Marks et al. (2012) Cell [GSE23943](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE23943) in E14-Tg2a mESCs cultured in serum and in 2i.

## Methods

For large downloads, we used Entrez direct + SRA prefetch. We first downloaded data in sra format and then used a parallelized version of fastq-dump to considerably speed up dumping. Of note, an Aspera client instance would be even faster.
Metadata in SOFT format were downloaded and mined to generate standard metadata files for our pipeline, adding SRA experiment and SRA run ids using an SRA to GEO lookup table. Raw files were concatenated by biological replicates and systematically renamed according to metadata.

## Data download and metadata mining

#### Joshi et al.

```{bash, eval = FALSE}
# prefetch data
esearch -db sra -query SRP062574 | efetch --format runinfo | cut -d ',' -f 1 | grep SRR | xargs prefetch -p 0.1 --max-size 500000000

# dump to fq
for file in *.sra; do parallel-fastq-dump --sra-id $file --threads 24 --outdir /home/f.comoglio/mydata/Projects/EPmatrix/Pub/GSE72164/ --split-3 --gzip; done; rm *.sra
```

#### Factor et al.

```{bash, eval = FALSE}
# prefetch data
esearch -db sra -query SRP041756 | efetch --format runinfo | cut -d ',' -f 1 | grep SRR | xargs prefetch -p 0.1 --max-size 500000000 &
esearch -db sra -query SRP041758 | efetch --format runinfo | cut -d ',' -f 1 | grep SRR | xargs prefetch -p 0.1 --max-size 500000000 &
esearch -db sra -query SRP041759 | efetch --format runinfo | cut -d ',' -f 1 | grep SRR | xargs prefetch -p 0.1 --max-size 500000000 &

# dump to fq
for file in *.sra; do parallel-fastq-dump --sra-id $file --threads 24 --outdir /home/f.comoglio/mydata/Projects/EPmatrix/Pub/GSE57409 --split-3 --gzip; done; rm *.sra
```

#### ENCODE

```{bash,eval =F}
# prefetch data
esearch -db sra -query SRX191012 | efetch --format runinfo | cut -d ',' -f 1 | grep SRR | xargs prefetch -p 0.1 --max-size 500000000

# dump to fq
for file in *.sra; do parallel-fastq-dump --sra-id $file --threads 24 --outdir /home/f.comoglio/mydata/Projects/EPmatrix/Pub/GSE37074 --split-3 --gzip; done; rm *.sra
```

#### Novo et al.

Significant CHi-C pairs in Seqmonk format were downloaded to `/home/f.comoglio/mydata/Projects/EPmatrix/Pub/GSE103053` and converted to a 12 column format to generate BED12 tracks and GenomicInteractions objects

```{bash, eval = FALSE}
# 6 cols --> 12 cols
# with process subst
for f in /home/f.comoglio/mydata/Projects/EPmatrix/Pub/GSE103053/Chic-seq/*.txt
do
  bn=$(basename $f)
  fn=${bn%%.*}
  paste <(awk 'NR % 2 == 1' $f) <(awk 'NR % 2 == 0' $f) | cut --complement -f7 > ${fn}_seqmonk_joined.txt
done
```

```{r tracks and gi objects, eval = FALSE}
path <- '/home/f.comoglio/mydata/Projects/EPmatrix/Pub/GSE103053/Chic-seq/'

# generate GenomicInteractions objects
gi.pchic.esc   <- getGI(paste0(path, 'GSE103053_ESC_complete_merge_step2_seqmonk_seqmonk_joined.txt'))
gi.pchic.episc <- getGI(paste0(path, 'GSE103053_EpiSC_merge_step2_seqmonk_seqmonk_joined.txt'))

# export GI as BED12
export.bed12(gi.pchic.esc, fn = paste0(path, 'Tracks/GSE103053_ESC_PCHiC.bed'), score = 'score')
export.bed12(gi.pchic.episc, fn = paste0(path, 'Tracks/GSE103053_EpiSC_PCHiC.bed'), score = 'score')
```

```{r save gi objects, eval = FALSE}
save(list = c('gi.pchic.esc', 'gi.pchic.episc'),
     file = here('../data/tad_deconstr_design/fc180315_epsure_tadec_pchic_mesc_novo_et_al.RData'))
```

#### Bonev et al.

TAD coordinates and A/B compartment assignment were downloaded to `/home/f.comoglio/mydata/Projects/EPmatrix/Pub/GSE96107`. 

#### Sherwood et al.

```{bash, eval = FALSE}
# prefetch data
esearch -db sra -query SRP034917 | efetch --format runinfo | cut -d ',' -f 1 | grep SRR | xargs prefetch -p 0.1 --max-size 500000000

# dump to fq
for file in *.sra; do parallel-fastq-dump --sra-id $file --threads 24 --outdir /home/f.comoglio/mydata/Projects/EPmatrix/Pub/GSE53776/ --split-3 --gzip; done; rm *.sra
```

#### Kagey et al.

```{bash, eval = FALSE}
# prefetch data
esearch -db sra -query SRP002778 | efetch --format runinfo | cut -d ',' -f 1 | grep SRR | xargs prefetch -p 0.1 --max-size 500000000

# dump to fq
for file in *.sra; do parallel-fastq-dump --sra-id $file --threads 24 --outdir /home/f.comoglio/mydata/Projects/EPmatrix/Pub/GSE22562/ --split-3 --gzip; done; rm *.sra
```

#### Whyte et al.

```{bash, eval = FALSE}
# prefetch data
esearch -db sra -query SRP018619 | efetch --format runinfo | cut -d ',' -f 1 | grep SRR | xargs prefetch -p 0.1 --max-size 500000000

# dump to fq
for file in *.sra; do parallel-fastq-dump --sra-id $file --threads 24 --outdir /home/f.comoglio/mydata/Projects/EPmatrix/Pub/GSE44286/ --split-3 --gzip; done; rm *.sra
```

#### Galonska et al.

```{bash, eval = FALSE}
# prefetch data
esearch -db sra -query SRP040666 | efetch --format runinfo | cut -d ',' -f 1 | grep SRR | xargs prefetch -p 0.1 --max-size 500000000

# dump to fq
for file in *.sra; do parallel-fastq-dump --sra-id $file --threads 24 --outdir /home/f.comoglio/mydata/Projects/EPmatrix/Pub/GSE56312/ --split-3 --gzip; done; rm *.sra
```

#### Chen et al. 

```{bash, eval = FALSE}
# prefetch data
esearch -db sra -query SRP000217 | efetch --format runinfo | cut -d ',' -f 1 | grep SRR | xargs prefetch -p 0.1 --max-size 500000000

# dump to fq
for file in *.sra; do parallel-fastq-dump --sra-id $file --threads 24 --outdir /home/f.comoglio/mydata/Projects/EPmatrix/Pub/GSE11431/ --split-3 --gzip; done; rm *.sra
```

### Marks et al.

```{bash, eval = FALSE}
# prefetch data
esearch -db sra -query SRP003412 | efetch --format runinfo | cut -d ',' -f 1 | grep SRR | xargs prefetch -p 0.1 --max-size 500000000

# dump to fq
for file in *.sra; do parallel-fastq-dump --sra-id $file --threads 24 --outdir /home/f.comoglio/mydata/Projects/EPmatrix/Pub/GSE23943/ --split-3 --gzip; done; rm *.sra
```

### Mining metadata

```{bash, eval = FALSE}
# download metadata
mkcdir /home/f.comoglio/mydata/Projects/EPmatrix/Pub/Metadata
wget ftp://ftp.ncbi.nlm.nih.gov/geo/series/GSE72nnn/GSE72164/soft/GSE72164_family.soft.gz
wget ftp://ftp.ncbi.nlm.nih.gov/geo/series/GSE57nnn/GSE57409/soft/GSE57409_family.soft.gz
wget ftp://ftp.ncbi.nlm.nih.gov/geo/series/GSE37nnn/GSE37074/soft/GSE37074_family.soft.gz
wget ftp://ftp.ncbi.nlm.nih.gov/geo/series/GSE53nnn/GSE53776/soft/GSE53776_family.soft.gz
wget ftp://ftp.ncbi.nlm.nih.gov/geo/series/GSE22nnn/GSE22562/soft/GSE22562_family.soft.gz
wget ftp://ftp.ncbi.nlm.nih.gov/geo/series/GSE44nnn/GSE44286/soft/GSE44286_family.soft.gz
wget ftp://ftp.ncbi.nlm.nih.gov/geo/series/GSE56nnn/GSE56312/soft/GSE56312_family.soft.gz
wget ftp://ftp.ncbi.nlm.nih.gov/geo/series/GSE11nnn/GSE11431/soft/GSE11431_family.soft.gz
wget ftp://ftp.ncbi.nlm.nih.gov/geo/series/GSE23nnn/GSE23943/soft/GSE23943_family.soft.gz
```

```{r load lookup table, eval = FALSE}
tib.lookup <- read_tsv('/home/f.comoglio/mydata/SRA/SRR_to_GSM.tab', col_names = FALSE) %>%
  mutate(sra.run = X1, geo.accession = X10, sra.experiment = X11) %>%
  select(sra.run, sra.experiment, geo.accession) %>%
  mutate(geo.accession = sapply(strsplit(geo.accession, '\\.'), '[', 1),
         geo.accession = sapply(strsplit(geo.accession, '_'), '[', 1))
```

```{r mine metadata, eval = FALSE}
# Joshi et al.
extractMeta(soft.fn = '/home/f.comoglio/mydata/Projects/EPmatrix/Pub/Metadata/GSE72164_family.soft.gz', 
            fq.dir = '/home/f.comoglio/mydata/Projects/EPmatrix/Pub/Raw/GSE72164', 
            tib.lookup = tib.lookup,
            raw.format = 'fastq.gz',
            meta.out.fn = '/home/f.comoglio/mydata/Projects/EPmatrix/Pub/Metadata/GSE72164_metadata.tsv', 
            meta.out.concat.fn = '/home/f.comoglio/mydata/Projects/EPmatrix/Pub/Metadata/GSE72164_metadata_concat.tsv')

# Factor et al.
extractMeta(soft.fn = '/home/f.comoglio/mydata/Projects/EPmatrix/Pub/Metadata/GSE57409_family.soft.gz', 
            fq.dir = '/home/f.comoglio/mydata/Projects/EPmatrix/Pub/GSE57409', 
            tib.lookup = tib.lookup,
            raw.format = 'fastq.gz',
            meta.out.fn = '/home/f.comoglio/mydata/Projects/EPmatrix/Pub/Metadata/GSE57409_metadata.tsv', 
            meta.out.concat.fn = '/home/f.comoglio/mydata/Projects/EPmatrix/Pub/Metadata/GSE57409_metadata_concat.tsv')

# ENCODE
# required manual curation as biological replicates not specified in metadata from ENCODE
# and metadata are for entire superseries (only DNase sample retained)
extractMeta(soft.fn = '/home/f.comoglio/mydata/Projects/EPmatrix/Pub/Metadata/GSE37074_family.soft.gz', 
            fq.dir = '/home/f.comoglio/mydata/Projects/EPmatrix/Pub/GSE37074', 
            tib.lookup = tib.lookup,
            raw.format = 'fastq.gz',
            meta.out.fn = '/home/f.comoglio/mydata/Projects/EPmatrix/Pub/Metadata/GSE37074_metadata.tsv', 
            meta.out.concat.fn = '/home/f.comoglio/mydata/Projects/EPmatrix/Pub/Metadata/GSE37074_metadata_concat.tsv')

# Sherwood et al.
extractMeta(soft.fn = '/home/f.comoglio/mydata/Projects/EPmatrix/Pub/Metadata/GSE53776_family.soft.gz', 
            fq.dir = '/home/f.comoglio/mydata/Projects/EPmatrix/Pub/GSE53776', 
            tib.lookup = tib.lookup,
            raw.format = 'fastq.gz',
            meta.out.fn = '/home/f.comoglio/mydata/Projects/EPmatrix/Pub/Metadata/GSE53776_metadata.tsv', 
            meta.out.concat.fn = '/home/f.comoglio/mydata/Projects/EPmatrix/Pub/Metadata/GSE53776_metadata_concat.tsv')

# Kagey et al.
# required manual curation as sample labels were not following good practices
extractMeta(soft.fn = '/home/f.comoglio/mydata/Projects/EPmatrix/Pub/Metadata/GSE22562_family.soft.gz', 
            fq.dir = '/home/f.comoglio/mydata/Projects/EPmatrix/Pub/GSE22562', 
            tib.lookup = tib.lookup,
            raw.format = 'fastq.gz',
            meta.out.fn = '/home/f.comoglio/mydata/Projects/EPmatrix/Pub/Metadata/GSE22562_metadata.tsv', 
            meta.out.concat.fn = '/home/f.comoglio/mydata/Projects/EPmatrix/Pub/Metadata/GSE22562_metadata_concat.tsv')

# Whyte et al.
# required manual curation as sample labels were not following good practices
extractMeta(soft.fn = '/home/f.comoglio/mydata/Projects/EPmatrix/Pub/Metadata/GSE44286_family.soft.gz', 
            fq.dir = '/home/f.comoglio/mydata/Projects/EPmatrix/Pub/GSE44286', 
            tib.lookup = tib.lookup,
            raw.format = 'fastq.gz',
            meta.out.fn = '/home/f.comoglio/mydata/Projects/EPmatrix/Pub/Metadata/GSE44286_metadata.tsv', 
            meta.out.concat.fn = '/home/f.comoglio/mydata/Projects/EPmatrix/Pub/Metadata/GSE44286_metadata_concat.tsv')

# Galonka et al.
# required manual curation as sample labels were not following good practices for biological replicates
extractMeta(soft.fn = '/home/f.comoglio/mydata/Projects/EPmatrix/Pub/Metadata/GSE56312_family.soft.gz', 
            fq.dir = '/home/f.comoglio/mydata/Projects/EPmatrix/Pub/GSE56312', 
            tib.lookup = tib.lookup,
            raw.format = 'fastq.gz',
            meta.out.fn = '/home/f.comoglio/mydata/Projects/EPmatrix/Pub/Metadata/GSE56312_metadata.tsv', 
            meta.out.concat.fn = '/home/f.comoglio/mydata/Projects/EPmatrix/Pub/Metadata/GSE56312_metadata_concat.tsv')

# Chen et al.
# required manual curation as SRR identifiers not present within metadata (data are from 2008)
extractMeta(soft.fn = '/home/f.comoglio/mydata/Projects/EPmatrix/Pub/Metadata/GSE11431_family.soft.gz', 
            fq.dir = '/home/f.comoglio/mydata/Projects/EPmatrix/Pub/GSE11431', 
            tib.lookup = tib.lookup,
            raw.format = 'fastq.gz',
            meta.out.fn = '/home/f.comoglio/mydata/Projects/EPmatrix/Pub/Metadata/GSE11431_metadata.tsv', 
            meta.out.concat.fn = '/home/f.comoglio/mydata/Projects/EPmatrix/Pub/Metadata/GSE11431_metadata_concat.tsv')

# Marks et al.
extractMeta(soft.fn = '/home/f.comoglio/mydata/Projects/EPmatrix/Pub/Metadata/GSE23943_family.soft.gz', 
            fq.dir = '/home/f.comoglio/mydata/Projects/EPmatrix/Pub/GSE23943', 
            tib.lookup = tib.lookup,
            raw.format = 'fastq.gz',
            meta.out.fn = '/home/f.comoglio/mydata/Projects/EPmatrix/Pub/Metadata/GSE23943_metadata.tsv', 
            meta.out.concat.fn = '/home/f.comoglio/mydata/Projects/EPmatrix/Pub/Metadata/GSE23943_metadata_concat.tsv')
```

### Concatenating and renaming files

```{bash, eval = FALSE}
# Joshi et al.
while read id runlist
do
  ofn="${id}.fastq.gz"
  cat ${runlist} > ${ofn}
done </home/f.comoglio/mydata/Projects/EPmatrix/Pub/Metadata/GSE72164_metadata_concat.tsv
rm SRR*

# Factor et al.
cd /home/f.comoglio/mydata/Projects/EPmatrix/Pub/GSE57409
while read id runlist
do
  ofn="${id}.fastq.gz"
  cat ${runlist} > ${ofn}
done </home/f.comoglio/mydata/Projects/EPmatrix/Pub/Metadata/GSE57409_metadata_concat.tsv
rm SRR*

# ENCODE
cd /home/f.comoglio/mydata/Projects/EPmatrix/Pub/GSE37074
while read id runlist
do
  ofn="${id}.fastq.gz"
  cat ${runlist} > ${ofn}
done </home/f.comoglio/mydata/Projects/EPmatrix/Pub/Metadata/GSE37074_metadata_concat.tsv
rm SRR*

# Sherwood et al.
cd /home/f.comoglio/mydata/Projects/EPmatrix/Pub/GSE53776
while read id runlist
do
  ofn="${id}.fastq.gz"
  cat ${runlist} > ${ofn}
done </home/f.comoglio/mydata/Projects/EPmatrix/Pub/Metadata/GSE53776_metadata_concat.tsv
rm SRR*

# Kagey et al.
cd /home/f.comoglio/mydata/Projects/EPmatrix/Pub/GSE22562
while read id runlist
do
  ofn="${id}.fastq.gz"
  cat ${runlist} > ${ofn}
done </home/f.comoglio/mydata/Projects/EPmatrix/Pub/Metadata/GSE22562_metadata_concat.tsv
rm SRR*

# Whyte et al.
cd /home/f.comoglio/mydata/Projects/EPmatrix/Pub/GSE44286
while read id runlist
do
  ofn="${id}.fastq.gz"
  cat ${runlist} > ${ofn}
done </home/f.comoglio/mydata/Projects/EPmatrix/Pub/Metadata/GSE44286_metadata_concat.tsv
rm SRR*

# Galonska et al.
cd /home/f.comoglio/mydata/Projects/EPmatrix/Pub/GSE56312
while read id runlist
do
  ofn="${id}.fastq.gz"
  cat ${runlist} > ${ofn}
done </home/f.comoglio/mydata/Projects/EPmatrix/Pub/Metadata/GSE56312_metadata_concat.tsv
rm SRR*

# Chen et al.
cd /home/f.comoglio/mydata/Projects/EPmatrix/Pub/GSE11431
while read id runlist
do
  ofn="${id}.fastq.gz"
  cat ${runlist} > ${ofn}
done </home/f.comoglio/mydata/Projects/EPmatrix/Pub/Metadata/GSE11431_metadata_concat.tsv
rm SRR*

# Marks et al.
cd /home/f.comoglio/mydata/Projects/EPmatrix/Pub/GSE23943
while read id runlist
do
  ofn="${id}.fastq.gz"
  cat ${runlist} > ${ofn}
done </home/f.comoglio/mydata/Projects/EPmatrix/Pub/Metadata/GSE23943_metadata_concat.tsv
rm SRR*
```

## UCSC session and track hubs

An up-to-date UCSC session with all processed data can be accessed at http://genome-euro.ucsc.edu/cgi-bin/hgTracks?hgS_doOtherUser=submit&hgS_otherUserName=Federico.comoglio&hgS_otherUserSessionName=01_TAD_deconstr_mESC

For records, the following track hubs are hosted on lovelace (/DATA/public):

https://rhpc.nki.nl/lovelace/Hubs/DNase_Joshi_et_al_2015/hub.txt
https://rhpc.nki.nl/lovelace/Hubs/ChIP_Joshi_et_al_2015/hub.txt
https://rhpc.nki.nl/lovelace/Hubs/DNase_JStam/hub.txt
https://rhpc.nki.nl/lovelace/Hubs/ChIP_Factor_et_al_2014/hub.txt
https://rhpc.nki.nl/lovelace/Hubs/DNase_Factor_et_al_2014/hub.txt
https://rhpc.nki.nl/lovelace/Hubs/DNase_Sherwood_et_al_2014/hub.txt
https://rhpc.nki.nl/lovelace/Hubs/ChIP_Kagey_et_al_2010/hub.txt
https://rhpc.nki.nl/lovelace/Hubs/ChIP_Whyte_et_al_2013/hub.txt
https://rhpc.nki.nl/lovelace/Hubs/ChIP_Galonska_et_al_2015/hub.txt
https://rhpc.nki.nl/lovelace/Hubs/ChIP_Chen_et_al_2008/hub.txt
https://rhpc.nki.nl/lovelace/Hubs/ChIP_Marks_et_al_2012/hub.txt
https://rhpc.nki.nl/lovelace/Hubs/RNAseq_Joshi_et_al_2015/hub.txt
https://rhpc.nki.nl/lovelace/Hubs/RNAseq_Factor_et_al_2014/hub.txt

## Data post-processing

### DHS CHi-C from Joshi et al.

Truncated fastq files and filtered BAM (intermediate HiCUP output files) were removed to free disk space. The remaining files were organized into subdirectories as follows.

```{bash, eval = FALSE}
resdir=/home/f.comoglio/Pub/GSE72164/Chic-seq/Results/HiCUP
rm $resdir/*trunc.fastq.gz
rm $resdir/*.filt.bam
rm -r $resdir/hicup_filter_ditag_rejects_13-01-58_21-03-2018

mkdir $resdir/Summary $resdir/Plots $resdir/Paired $resdir/Valid
mv $resdir/*summary* $resdir/Summary
mv $resdir/*.svg $resdir/Plots
mv $resdir/*.pair.bam $resdir/Paired
mv $resdir/*.hicup*.bam $resdir/Valid
```

Merged BAM files containing the union of valid di-tags from both biological replicates were generated (without deduplication). These were sorted, indexed and tabix indexed.

```{bash, eval = FALSE}
samtools cat -o $resdir/Valid/GSM1856416_GSM1856417_CHi-C_mESCs_serum_r1_2.hicup.bam $resdir/Valid/GSM1856416_CHi-C_mESCs_serum_BR1_r1_2.hicup.bam $resdir/Valid/GSM1856417_CHi-C_mESCs_serum_BR2_r1_2.hicup.bam

samtools cat -o $resdir/Valid/GSM1856418_GSM1856419_CHi-C_mESCs_2i_r1_2.hicup.bam $resdir/Valid/GSM1856418_CHi-C_mESCs_2i_BR1_r1_2.hicup.bam $resdir/Valid/GSM1856419_CHi-C_mESCs_2i_BR2_r1_2.hicup.bam

# sort merged BAM (indexing)
sambamba sort --nthreads=16 --out=$resdir/Valid/GSM1856416_GSM1856417_CHi-C_mESCs_serum_r1_2.hicup_sorted.bam $resdir/Valid/GSM1856416_GSM1856417_CHi-C_mESCs_serum_r1_2.hicup.bam
sambamba sort --nthreads=16 --out=$resdir/Valid/GSM1856418_GSM1856419_CHi-C_mESCs_2i_r1_2.hicup_sorted.bam $resdir/Valid/GSM1856418_GSM1856419_CHi-C_mESCs_2i_r1_2.hicup.bam

# generate tabix files
python /home/f.comoglio/gitlab/gen-scripts/v4c/hicup_to_v4c.py $resdir/Valid/GSM1856416_GSM1856417_CHi-C_mESCs_serum_r1_2.hicup_sorted.bam
python /home/f.comoglio/gitlab/gen-scripts/v4c/hicup_to_v4c.py $resdir/Valid/GSM1856418_GSM1856419_CHi-C_mESCs_2i_r1_2.hicup_sorted.bam
```

### DHS CHi-C baits from Joshi et al.

For DHS CHi-C data, the coordinates of the baited regions were not deposited by the authors. However, in the methods they state that "To improve quality, the whole genome was windowed into 5 DpnII site tiles". Therefore, we extracted these baited regions from the combined CHiCAGO significant interactions file (mm9) at  [GSE72164](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE72164).

```{bash, eval = FALSE}
odir=/home/f.comoglio/Pub/GSE72164/Chic-seq/Author_deposited
cut -f1-3 $odir/GSE72164_Hi-C_Merged_interactions_intervals.txt | sort -u | tail -n +2 > $odir/GSE72164_DHS_CHiC_baits_mm9.bed
```

and lifted over to mm10

```{r dhs chic bait coord, eval = FALSE}
dhs.chic.baits.mm10 <- liftOver(peakFile = '/home/f.comoglio/Pub/GSE72164/Chic-seq/Author_deposited/GSE72164_DHS_CHiC_baits_mm9.bed', 
                                chainFile = '/home/f.comoglio/mydata/Annotations/mm9ToMm10.over.chain')

# export to bed track
# bed is 0-based
tmp <- dhs.chic.baits.mm10
start(tmp) <- start(tmp) + 1
rtracklayer::export(tmp, con = '/home/f.comoglio/Pub/GSE72164/Chic-seq/Author_deposited/GSE72164_DHS_CHiC_baits_mm10.bed', format = 'bed')

# save object
save(list = c('dhs.chic.baits.mm10'),
     file = here('../data/tad_deconstr_design/fc180328_epsure_tadec_dhs_chic_bait_coord_joshi_et_al.RData'))
```

This bait could be used at a later stage to call significant interactions with CHiCAGO. For the current analyses, the combined CHiCAGO significant interactions file was lifted over to mm10 as follows.

```{sh liftover dhs chic part 1, eval = FALSE}
odir=/home/f.comoglio/Pub/GSE72164/Chic-seq/Author_deposited

# 0. check that file contains main chr only
cut -f 1 $odir/GSE72164_Hi-C_Merged_interactions_intervals.txt | sort | uniq

# 1. remove header and add unique row identifier
cat $odir/GSE72164_Hi-C_Merged_interactions_intervals.txt | tail -n +2 | awk -v OFS='\t' '{print $0,NR}' > $odir/GSE72164_Hi-C_Merged_interactions_intervals_with_row_id.ibed

# 2. extract bait columns and prey columns + row id
cut -f 1,2,3,23 $odir/GSE72164_Hi-C_Merged_interactions_intervals_with_row_id.ibed > bait_coord_mm9.bed
cut -f 4,5,6,23 $odir/GSE72164_Hi-C_Merged_interactions_intervals_with_row_id.ibed > prey_coord_mm9.bed
cut -f 4-23 $odir/GSE72164_Hi-C_Merged_interactions_intervals_with_row_id.ibed > prey_coord_with_scores_mm9.bed

# 3. liftOver to mm10
liftOver bait_coord_mm9.bed /home/f.comoglio/mydata/Annotations/mm9ToMm10.over.chain bait_coord_mm10.bed unmapped_bait
liftOver prey_coord_mm9.bed /home/f.comoglio/mydata/Annotations/mm9ToMm10.over.chain prey_coord_mm10.bed unmapped_prey

# check how many rows mapped
cut -f 4 bait_coord_mm10.bed | uniq | wc -l
cut -f 4 prey_coord_mm10.bed | uniq | wc -l
```

```{r liftover dhs chic part 2, eval = FALSE}
# liftOver expects additional fields to be integers and hence it truncates scores
# add score to prey file, then reconstruct interaction file

df_bait                 <- read.table('/home/f.comoglio/Pub/GSE72164/Chic-seq/Author_deposited/bait_coord_mm10.bed', as.is = TRUE)
colnames(df_bait)       <- c('baitChr',	'baitStart',	'baitEnd', 'id')
df_prey                 <- read.table('/home/f.comoglio/Pub/GSE72164/Chic-seq/Author_deposited/prey_coord_mm10.bed', as.is = TRUE)
colnames(df_prey)       <- c('oeChr',	'oeStart', 'oeEnd', 'id')
df_prey_score           <- read.table('/home/f.comoglio/Pub/GSE72164/Chic-seq/Author_deposited/prey_coord_with_scores_mm9.bed', as.is = TRUE)
colnames(df_prey_score) <- c('oeChr',	'oeStart',	'oeEnd',	'CapC_serum_merged',	'CapC_2i_merged',	'CapC_serum_6608',	'CapC_serum_7940',	'CapC_2i_6609',	'CapC_2i_7941',	'CapC_2i_d1_7942',	'CapC_2i_d3_7924',	'CapC_Eed_8704',	'CapC_serum_6608.chinput',	'CapC_serum_7940.chinput',	'CapC_2i_6609.chinput',	'CapC_2i_7941.chinput',	'CapC_2i_d1_7942.chinput',	'CapC_2i_d3_7924.chinput',	'CapC_Eed_8704.chinput', 'id')

df_int <- left_join(df_bait, df_prey, by = 'id') %>%
  left_join(dplyr::select(df_prey_score, -c(oeChr, oeStart, oeEnd)), by = 'id') %>%
  filter(!is.na(baitChr), !is.na(baitStart), !is.na(baitEnd), !is.na(oeChr), !is.na(oeStart), !is.na(oeEnd)) %>%
  dplyr::select(-id, -contains('chinput'))

# write to file
write.table(df_int, file = '/home/f.comoglio/Pub/GSE72164/Chic-seq/Author_deposited/GSE72164_Hi-C_Merged_interactions_intervals_mm10.txt', 
            quote = FALSE, sep = '\t', row.names = FALSE, col.names = TRUE)

# get GInteraction object
gi_dhs_chic_joshi_2i    <- get_gint_joshi(df_int, score_col_name = 'CapC_2i_merged') 
gi_dhs_chic_joshi_serum <- get_gint_joshi(df_int, score_col_name = 'CapC_serum_merged') 

# export GI as BED12
export.bed12(gi_dhs_chic_joshi_2i, 
             fn = '/home/f.comoglio/mydata/Projects/EPmatrix/Pub/GSE72164/Chic-seq/Tracks/GSE72164_DHS_CHiC_Joshi_mESC_2i.bed', score = 'score')
export.bed12(gi_dhs_chic_joshi_serum, 
             fn = '/home/f.comoglio/mydata/Projects/EPmatrix/Pub/GSE72164/Chic-seq/Tracks/GSE72164_DHS_CHiC_Joshi_mESC_serum.bed', score = 'score')

# load all bait coordinates
df <- read.table('/home/f.comoglio/Pub/GSE72164/Chic-seq/Author_deposited/GSE72164_DHS_CHiC_baits_mm10.bed', as.is = TRUE)[, 1:3]
colnames(df) <- c('seqnames', 'start', 'end')
gr_dhs_baits <- as(df, 'GRanges')
```

```{r save gi objects joshi, eval = FALSE}
save(list = c('gi_dhs_chic_joshi_2i', 'gi_dhs_chic_joshi_serum', 'gr_dhs_baits'),
     file = here('../data/tad_deconstr_design/fc180904_epsure_tadec_dhs_chic_mesc_joshi_et_al.RData'))
```

### TAD coordinates from Bonev et al.

A BED file and a GRanges object with TAD coordinates were generated as follows.

```{r tad coord, eval = FALSE}
gr.tad.esc <- readxl::read_xlsx('/home/f.comoglio/mydata/Projects/EPmatrix/Pub/GSE96107/Hi-c/1-s2.0-S0092867417311376-mmc2.xlsx') %>%
  dplyr::rename(seqnames = chrom) %>%
  as('GRanges')

# export to bed track
# bed is 0-based
tmp <- gr.tad.esc
start(tmp) <- start(tmp) + 1
rtracklayer::export(tmp, con = '/home/f.comoglio/mydata/Projects/EPmatrix/Pub/GSE96107/Hi-c/Tracks/GSE96107_mESC_TAD_coord.bed', format = 'bed')

# save object
save(list = c('gr.tad.esc'),
     file = here('../data/tad_deconstr_design/fc180319_epsure_tadec_tad_coord_mesc_bonev_et_al.RData'))
```

### DHS mining (all DNase experiments)

We generated two sets of DHS while optimizing DAT:

* Relaxed, all detected DHSs
* Stringent, only DHSs with p-value for local signal enrichment over estimated background of <= 1e-10

#### Relaxed set

```{r dhs mining}
dhs.2i           <- getDHS('/home/f.comoglio/Pub/GSE72164/Dnase-seq/Results/Peaks/peaks_GSM1856456_DNaseI_2i.bed')
# dhs.serum.joshi  <- getDHS('/home/f.comoglio/Pub/GSE72164/Dnase-seq/Results/Peaks/peaks_GSM1856455_DNaseI_serum.bed')

dhs.serum.stam.1 <- getDHS('/home/f.comoglio/Pub/GSE37074/Dnase-seq/Results/Peaks/peaks_GSM1014154_UW_DnaseSeq_ES-E14_E0_129_Ola_BR1.bed')
dhs.serum.stam.2 <- getDHS('/home/f.comoglio/Pub/GSE37074/Dnase-seq/Results/Peaks/peaks_GSM1014154_UW_DnaseSeq_ES-E14_E0_129_Ola_BR2.bed')

# combine replicates
dhs.serum        <- combineRepl(dhs.serum.stam.1, dhs.serum.stam.2)

dhs.mesendo      <- getDHS('/home/f.comoglio/Pub/GSE53776/Dnase-seq/Results/Peaks/peaks_GSM1300603_DNase_mES_Day3_Mesendoderm.bed')
dhs.meso         <- getDHS('/home/f.comoglio/Pub/GSE53776/Dnase-seq/Results/Peaks/peaks_GSM1300604_DNase_mES_Day5_Mesoderm.bed')
dhs.endo         <- getDHS('/home/f.comoglio/Pub/GSE53776/Dnase-seq/Results/Peaks/peaks_GSM1300605_DNase_mES_Day5_Endoderm.bed')
#dhs.intest       <- getDHS('/home/f.comoglio/Pub/GSE53776/Dnase-seq/Results/Peaks/peaks_GSM1300606_DNase_mES_Day6_Intestinal_Endoderm.bed')
#dhs.prepancr     <- getDHS('/home/f.comoglio/Pub/GSE53776/Dnase-seq/Results/Peaks/peaks_GSM1300607_DNase_mES_Day6_Prepancreatic_Endoderm.bed')

# select de novo DHS not detected in 2i/serum
# useful for visualization
dhs.devel <- c(dhs.mesendo, dhs.meso, dhs.endo) %>%
  GenomicRanges::reduce()
dhs.late  <- subsetByOverlaps(dhs.devel, union(dhs.2i, dhs.serum[, 0]), invert = TRUE)
```

```{r dhs peak export, echo = FALSE, eval = FALSE}
# export dhs coordinates to bed track (0-based)
tmp <- dhs.2i; start(tmp) <- start(tmp) + 1
rtracklayer::export(tmp, con = '/home/f.comoglio/Pub/GSE72164/Dnase-seq/Results/Peaks/peaks_GSM1856456_DNaseI_2i_postproc.bed', format = 'bed')
tmp <- dhs.serum.joshi; start(tmp) <- start(tmp) + 1
rtracklayer::export(tmp, con = '/home/f.comoglio/Pub/GSE72164/Dnase-seq/Results/Peaks/peaks_GSM1856455_DNaseI_serum_postproc.bed', format = 'bed')
tmp <- dhs.serum; start(tmp) <- start(tmp) + 1
rtracklayer::export(tmp, con = '/home/f.comoglio/Pub/GSE37074/Dnase-seq/Results/Peaks/peaks_GSM1014154_UW_DnaseSeq_ES-E14_E0_129_Ola_postproc.bed', format = 'bed')
tmp <- dhs.late; start(tmp) <- start(tmp) + 1
rtracklayer::export(tmp, con = '/home/f.comoglio/Pub/GSE53776/Dnase-seq/Results/Peaks/peaks_GSM1300603_DNase_devel_late_postproc.bed', format = 'bed')
```

```{r upset plot dhs, fig.width = 8}
dhs.grl <- GRangesList('2i' = dhs.2i, 'serum' = dhs.serum[, 0], 'mesendo' = dhs.mesendo, 
                       'meso' = dhs.meso, 'endo' = dhs.endo)
master.set <- reduce(unlist(dhs.grl))

#compute olap --> format for upset (data frame)
olap.df    <- sapply(dhs.grl, function(gr) master.set %over% gr %>% as.numeric) %>% as.data.frame

upset(data = olap.df,
      nsets = ncol(olap.df),
      keep.order = TRUE,
      order.by = c('freq'),
      sets = c('2i', 'serum', 'mesendo', 'meso', 'endo'),
      nintersects = 20,
      queries = list(list(query = intersects, params = list('2i', 'serum'), color = "orange", active = TRUE)))
```

```{r save dhs objects, eval = FALSE}
# save object
save(list = c('dhs.2i', 'dhs.serum', 'dhs.mesendo', 'dhs.endo', 'dhs.meso'),
     file = here('../data/tad_deconstr_design/fc180329_epsure_tadec_dhs_coord.RData'))
```

#### Stringent set

```{sh dhs mining stringent, eval = FALSE}
# filter by p-val <= 1e-10
tail -n +36 /home/f.comoglio/Pub/GSE72164/Dnase-seq/Results/Peaks/peaks_GSM1856456_DNaseI_2i.txt | awk '{ if ($10 <= 1e-10) { print } }' | cut -f 2,3,4 | sort -k1,1 > /home/f.comoglio/Pub/GSE72164/Dnase-seq/Results/Peaks/peaks_GSM1856456_DNaseI_2i_stringent.bed

tail -n +36 /home/f.comoglio/Pub/GSE37074/Dnase-seq/Results/Peaks/peaks_GSM1014154_UW_DnaseSeq_ES-E14_E0_129_Ola_BR1.txt | awk '{ if ($10 <= 1e-10) { print } }' | cut -f 2,3,4 | sort -k1,1 > /home/f.comoglio/Pub/GSE37074/Dnase-seq/Results/Peaks/peaks_GSM1014154_UW_DnaseSeq_ES-E14_E0_129_Ola_BR1_stringent.bed

tail -n +36 /home/f.comoglio/Pub/GSE37074/Dnase-seq/Results/Peaks/peaks_GSM1014154_UW_DnaseSeq_ES-E14_E0_129_Ola_BR2.txt | awk '{ if ($10 <= 1e-10) { print } }' | cut -f 2,3,4 | sort -k1,1 >  /home/f.comoglio/Pub/GSE37074/Dnase-seq/Results/Peaks/peaks_GSM1014154_UW_DnaseSeq_ES-E14_E0_129_Ola_BR2_stringent.bed
```

```{r dhs mining stringent robj, eval = FALSE}
dhs.2i           <- getDHS('/home/f.comoglio/Pub/GSE72164/Dnase-seq/Results/Peaks/peaks_GSM1856456_DNaseI_2i_stringent.bed', header.size = 0)
dhs.serum.stam.1 <- getDHS('/home/f.comoglio/Pub/GSE37074/Dnase-seq/Results/Peaks/peaks_GSM1014154_UW_DnaseSeq_ES-E14_E0_129_Ola_BR1_stringent.bed', header.size = 0)
dhs.serum.stam.2 <- getDHS('/home/f.comoglio/Pub/GSE37074/Dnase-seq/Results/Peaks/peaks_GSM1014154_UW_DnaseSeq_ES-E14_E0_129_Ola_BR2_stringent.bed', header.size = 0)

# combine replicates
dhs.serum        <- combineRepl(dhs.serum.stam.1, dhs.serum.stam.2)
```

```{r dhs peak export stringent, echo = FALSE, eval = FALSE}
# export dhs coordinates to bed track (0-based)
tmp <- dhs.2i; start(tmp) <- start(tmp) + 1
rtracklayer::export(tmp, con = '/home/f.comoglio/Pub/GSE72164/Dnase-seq/Results/Peaks/peaks_GSM1856456_DNaseI_2i_postproc_stringent.bed', format = 'bed')
tmp <- dhs.serum; start(tmp) <- start(tmp) + 1
rtracklayer::export(tmp, con = '/home/f.comoglio/Pub/GSE37074/Dnase-seq/Results/Peaks/peaks_GSM1014154_UW_DnaseSeq_ES-E14_E0_129_Ola_postproc_stringent.bed', format = 'bed')
```

```{r save dhs objects stringent, eval = FALSE}
# save object
save(list = c('dhs.2i', 'dhs.serum'),
     file = here('../data/tad_deconstr_design/fc180410_epsure_tadec_dhs_coord_stringent.RData'))
```


### ChIP-seq peak set (Chen et al.)

```{sh, eval = FALSE}
# peak calling
cd /home/f.comoglio/Pub/GSE11431/Chip-seq/Results/Homer

batchParallel.pl findPeaks homer.peaks -style factor -i GSM288358_ES_GFP -f GSM288345_ES_Nanog GSM288347_ES_Sox2 GSM288349_ES_E2f1 GSM288351_ES_CTCF GSM288353_ES_STAT3 GSM288355_ES_Esrrb  GSM288357_ES_n-Myc GSM288359_ES_p300 GSM288346_ES_Oct4 GSM288348_ES_Smad1 GSM288350_ES_Tcfcp2I1 GSM288352_ES_Zfx GSM288354_ES_Klf4 GSM288356_ES_c-Myc GSM288360_ES_Suz12

for f in *.peaks; do
 trackname=${f%%.*} # take prefix
 pos2bed.pl -bed -track $trackname $f
done

mkdir Peaks
mv *.peaks *.bed Peaks/

# peaks are in /home/f.comoglio/Pub/GSE11431/Chip-seq/Results/Homer/Peaks
```

```{r chipseq peaks to grl, eval = FALSE}
fn_vec <- list.files('/home/f.comoglio/Pub/GSE11431/Chip-seq/Results/Homer/Peaks', pattern = '*.bed', full.names = TRUE)

tf_names <- str_remove(fn_vec, '.*_') %>%
  str_remove('.bed') %>%
  str_replace('-', '') %>%
  toupper() %>%
  paste0('_ChIP')

grl_tf_chip_chen <- purrr::map(.x = fn_vec, .f = function(fn) homer_bed_to_gr(fn)) %>%
  GRangesList

names(grl_tf_chip_chen) <- tf_names
````

```{r save chipseq peaks objects, eval = FALSE}
# save object
save(list = c('grl_tf_chip_chen'),
     file = here('../data/tad_deconstr_design/fc180924_chipseq_chen_et_al_grl.RData'))
```


## Gencode annotations - R objects

```{r make gencode db, eval = FALSE}
# make gencode db from GTF file
gencode.db <- makeTxDbFromGFF(file = '~/mydata/Annotations/Genomes/mm10_gencode_v_m16/gencode.vM16.annotation.gtf', 
                              format = 'gtf', 
                              organism = 'Mus musculus', 
                              dataSource = 'gencode.vM16')

# save db
saveDb(gencode.db, file = '~/mydata/Annotations/mm10_gencode_v_m16.sqlite')

# extract gene promoters --> resize to TSS
#gencode.tss <- gencode.db %>%
#  genes %>%
#  promoters(0, 1) %>%
#  unique

#save(list = c('gencode.tss'),
#     file = here('../data/tad_deconstr_design/fc180402_epsure_tadec_gencode_tss.RData'))

# generate lookup tibble (gencode id to gene symbols)
# filter for protein coding genes and lincRNAs
gtf <- readGFF('~/mydata/Annotations/Genomes/mm10_gencode_v_m16/gencode.vM16.annotation.gtf')
gtf %<>% filter(type == 'gene', 
                gene_type %in% c('protein_coding', 'lincRNA'))

tib.gencode.lookup <- tibble(gene_id = gtf$gene_id, 
                             gene_name = gtf$gene_name, 
                             seqnames = gtf$seqid, 
                             start = gtf$start, 
                             end = gtf$end,
                             strand = gtf$strand) %>%
  distinct

# save tibble
save(list = c('tib.gencode.lookup'),
     file = here('../data/tad_deconstr_design/fc180402_epsure_tadec_gencode_lookup_tibble_pcod_lnc.RData'))

# extract protein coding/lincRNA gene promoters --> resize to TSS
pcod.linc.genes <- tib.gencode.lookup %>% 
  as('GRanges')

# extract gene promoters --> resize to TSS
pcod.linc.tss <- pcod.linc.genes %>%
  promoters(0, 1) %>%
  unique

save(list = c('pcod.linc.tss'),
     file = here('../data/tad_deconstr_design/fc180404_epsure_tadec_gencode_pcod_linc_tss.RData'))
```


## Gene expression tables

Gene expression count tables were generated from RNA-seq data by Joshi et al. (2i and serum). Two types of normalized counts were considered: rlog and rpkm.

```{r generate expression tables joshi}
# load DB
gencode_db <- loadDb(file = '~/mydata/Annotations/mm10_gencode_v_m16.sqlite')

# load protein coding gene annotation
load(here('../data/tad_deconstr_design/fc180402_epsure_tadec_gencode_lookup_tibble_pcod_lnc.RData'))
tib.gencode.lookup

exons_by_gene <- exonsBy(gencode_db, 'gene')

# define conditions
condition <- factor(c(rep('serum', 2), rep('twoi', 2)))
col_data  <- data.frame(condition = condition)

# load counts
counts    <- read.table('/home/f.comoglio/mydata/Projects/EPmatrix/Pub/GSE72164/Rna-seq/Results/Star/count_table.tsv', row.names = 1, header = TRUE)
colnames(counts)   <- paste0(condition, '_', 1:2)
rownames(col_data) <- colnames(counts)

# order according to gene ID in counts
exons_by_gene <- exons_by_gene[rownames(counts), ]

# create DESeq object
dds <- DESeqDataSetFromMatrix(countData = counts, 
                              colData   = col_data, 
                              design    = ~condition, 
                              rowRanges = exons_by_gene)

# normalize to rlog
rlog <- rlog(dds, blind = TRUE)	
rlog <- assay(rlog)

# normalize to rpkm
fpkm <- fpkm(dds, robust = TRUE)	

# format normalized counts
tib_rlog_joshi <- as.tibble(rlog, rownames = 'gene_id')
tib_fpkm_joshi <- as.tibble(fpkm, rownames = 'gene_id')

# add metadata
tib_rlog_joshi <- tib_rlog_joshi %>%
  left_join(tib.gencode.lookup, by = 'gene_id') %>%
  dplyr::select(gene_id, symbol = gene_name, seqnames, start, end, strand, everything()) %>% 
  mutate(symbol = str_to_upper(symbol)) %>%
  filter(!is.na(symbol))

tib_fpkm_joshi <- tib_fpkm_joshi %>%
  left_join(tib.gencode.lookup, by = 'gene_id') %>%
  dplyr::select(gene_id, symbol = gene_name, seqnames, start, end, strand, everything()) %>% 
  mutate(symbol = str_to_upper(symbol)) %>%
  filter(!is.na(symbol))
```

```{r save expression tables joshi}
save(list = c('tib_rlog_joshi', 'tib_fpkm_joshi'), file = here('../data/tad_deconstr_design/fc181121_epsure_tadec_rnaseq_expr_joshi.RData'))
```


## SessionInfo

```{r}
devtools::session_info()
```
