---
title: "MMA20201010_Tfcp2l1_Raw_to_Processed_V3pipeline_V4Norm"
author: "MMA"
date: "10/11/2020"
output:
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
editor_options: 
  chunk_output_type: inline
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo       = TRUE, 
                      message    = FALSE, 
                      warning    = FALSE, 
                      cache      = TRUE,
                      autodep    = TRUE,
                      fig.align  = 'center', 
                      fig.width  = 6, 
                      fig.height = 6)
```

```{r libs and fun}
library(tidyverse)
library(here)
library(magrittr)
library(reshape2)
library(knitr)
library(gridExtra)
library(pander)
library(cowplot)
library(circlize)
library(datteRo)
library(GGally)
library(ggplot2)
library(ggpubr)

```

---
# Intro
This is an update on `MMA190301_epsure_mesc_dat_Tfcp2l1_raw_to_coop_dattero.Rmd`. Data preprocessing has been updated to V3 epsure_pipeline, for both ecn and iPCR data. Here we update the data postprocessing to V4.

## Setup

8 cDNA and 2 pDNA samples were generated using EP014 (Tfcp2l1 Basal) and EP015 (Tfcp2l1 Combinatorial) libraries. The pDNA samples come from the libraries mix that was used to generate all the cDNA samples. 3 cDNA samples correspond to biological replicates in the 2i+lif culturing conditions and 5 correspond to biological replicates generated in the 2i-lif(Ml or minus_lif) culturing conditions. 3 of these 5 ssamples are biological replicates collected 24 hours after transfection as in the +lif condition. The other 2 samples are the same 2nd and 3rd replicates but collected 48hours after transfection.


**Notes**: for details on iPCR data analysis, see `MMA181130_epsure_mesc_dat_Tfcp2l1_raw_to_coop_dattero.Rmd`. iPCR data for the Tfcp2l1 libraries were analyzed using the ```ipcr.wf``` workflow of the EP-SuRE Snakemake pipeline v3 (see GitLab).

```{sh, eval = FALSE}
snakemake --cores 24 --use-conda --snakefile ~/mydata/GitLab/epmatrix/epsure_pipeline/v2/workflows/ecn.wf --configfile ~/mydata/GitLab/epmatrix/epsure_pipeline/v3/config/MMA20200629_ecn.yaml
```

Please refer to the same file for details on ECN (expression and copy number) data analysis.


## iPCR - Processing

### Basal library

```{r process basal lib}
# load basal library
tib_basal <- read_ipcr('/DATA/projects/epmatrix/mouse/DAT_EP014_EP015/ipcr/Results_v3/ipcr_E43_EP014_dat_tfcp2l1_basal_miseq_br1_unique_barcodes_frag_id_merged.tsv')
##Warning, this path is for my direct link to the shared folder, I could not make it work otherwise
# remove heterotypic

#store unfiltered version
tib_basal_unf<-tib_basal
#remove heterotypic combinations
tib_basal <- remove_heterotypic(tib_basal)

# fragment representation
count_fragments(tib_basal)
count_elements(tib_basal)
count_bc_reads(tib_basal)
```

```{r plot barcode count distribution, fig.width = 8, fig.height = 8}
plot_bc_distribution(tib_basal, type = 'basal')
```

```{r test orientation bias, fig.width = 4, fig.height = 4}
#plot_orientation_bias(tib_basal)
```

### Combinatorial library

```{r process comb lib}
# load combinatorial library
tib_comb <- read_ipcr('/DATA/projects/epmatrix/mouse/DAT_EP014_EP015/ipcr/Results_v3/ipcr_E43_EP015_dat_tfcp2l1_combinatorial_miseq_br1_combined_unique_barcodes_frag_id_merged.tsv')
#store unfiltered version
tib_comb_unf<-tib_comb
# count combinations
count_combinations(tib_comb, tib_basal)

#put appart empty fragments
tib_comb_empty<-tib_comb%>% filter(frag1=="eempty")

# mark and remove basal fragments from comb library. NOTE: It also removes empty fragments, that is why we put them appart before and bring them back afterwards.
tib_comb_het <- mark_basal(tib_comb) %>%
  filter(!with_basal)

# bring back empty elements
tib_comb_hetempty<-bind_rows(tib_comb_het,tib_comb_empty)

```

```{r comb lib - enhancer promoter representation, fig.width = 8, fig.height = 8}
tib_tmp <- count_combination_type(tib_comb_hetempty)
plot_combination_type(tib_tmp)

# summarize by combinations
#tib_comb_smrzd <- summarize_combinations(tib_comb)
#plot_bc_distribution(tib_comb_smrzd, type = 'comb')
```

```{r coop lib - identify hits with most combinations}
get_most_represented(tib_comb_hetempty)
get_most_represented(tib_comb_hetempty, prom_only = TRUE)
```


### Visualizing combinations

#### Binary matrix

```{r comb lib - heatmap represented comb, fig.width = 8, fig.height = 8}
# binary representation of fragment combinations (undirectional)
plot_ipcr_matrix(tib_comb_hetempty, type = 'binary', lab_frac = 0.5)
```

```{r comb lib - heatmap orient, fig.width = 10, fig.height = 8.5}
# number of orientations per combination
plot_ipcr_matrix(tib_comb_hetempty, type = 'orientations', lab_frac = 0.5)

# number of barcodes per combination
plot_ipcr_matrix(tib_comb_hetempty, type = 'barcodes', lab_frac = 0.5)
```


### get list of unique barcodes
```{r}
#merge all to check proper barcode clashing
all_ipcr_merged<-bind_rows(tib_basal_unf,tib_comb_unf)
# get unique barcodes before filtering
all_ipcr_merged_unique<-all_ipcr_merged %>%
  ungroup()%>%
  nest(-barcode ) %>% 
  mutate(n = map_dbl(data, nrow))%>%
  filter(n==1)%>%
  unnest()%>% 
  select(-n)
```

## Combining expression and copy number data

pDNA counts (copy number) are common for all of the experiments (p1+p2).
There are 3 cDNA biological replicates for +lif and -lif, and 2 biological replicates for -lif 48hours which are connected to 2 of the bilogical replicates of the -lif condition.
 

```{r batch2 merge ipcr and ecn data}
# read expression data (cdna)
tib_expr_br1 <- read_ecn('/DATA/projects/epmatrix/mouse/DAT_EP014_EP015/ecn/Results/cdna_E56_dat_br1_barcode_counts.tsv', col_lab = 'cdna_counts_br1')
tib_expr_br2 <- read_ecn('/DATA/projects/epmatrix/mouse/DAT_EP014_EP015/ecn/Results/cdna_E56_dat_br2_barcode_counts.tsv', col_lab = 'cdna_counts_br2')
tib_expr_br3 <- read_ecn('/DATA/projects/epmatrix/mouse/DAT_EP014_EP015/ecn/Results/cdna_E56_dat_br3_barcode_counts.tsv', col_lab = 'cdna_counts_br3')

# read copy number data (pdna)
tib_cn_br1 <- read_ecn('/DATA/projects/epmatrix/mouse/DAT_EP014_EP015/ecn/Results/pdna_E56_dat_br1_barcode_counts.tsv', col_lab = 'pdna_counts_br1')
tib_cn_br2 <- read_ecn('/DATA/projects/epmatrix/mouse/DAT_EP014_EP015/ecn/Results/pdna_E56_dat_br2_barcode_counts.tsv', col_lab = 'pdna_counts_br2')

# merge datasets
tib_basal_merged <- merge_ipcr_ecn(tib           = tib_basal, 
                                      cn_tib_list   = list(tib_cn_br1, tib_cn_br2),
                                      expr_tib_list = list(tib_expr_br1, tib_expr_br2, tib_expr_br3))

tib_comb_merged <- merge_ipcr_ecn(tib           = tib_comb_hetempty, 
                                     cn_tib_list   = list(tib_cn_br1, tib_cn_br2),
                                     expr_tib_list = list(tib_expr_br1, tib_expr_br2, tib_expr_br3))

get_barcode_summary(tib_basal, tib_comb, tib_basal_merged, tib_comb_merged)


```

```{r}
tib_comb_merged%>%
  filter(frag1=="eempty")
```


### Merging Comb and basal datasets, filter uniquely assigned barcodes
```{r}
#Plus lif samples
tib_basal_merged_idlib<-tib_basal_merged %>% 
  mutate(id=paste0(id, "_basal"))

tib_comb_merged_idlib<-tib_comb_merged %>% 
  mutate(id=paste0(id, "_comb"))


# v3_b1_v4 stands for v3 EP-sure pipeline, batch 1 expression data generation (only one batch for Tfcp2l1), v4 data normalisation.

tib_all_PlusLif_v3_b1_v4<-bind_rows(tib_basal_merged_idlib,tib_comb_merged_idlib)

#new barcode clashing removal
# This step is now crucial because there is no barcode clashing removal in the pipeline
#tib_all_NCklf2nanog_v3_b3_v4_unique <- tib_all_NCklf2nanog_v3_b3_v4 %>% ungroup()%>%
#  nest(-barcode ) %>% 
#  mutate(n = map_dbl(data, nrow))%>%
#  filter(n==1)%>%
#  unnest()%>% 
#  select(-n)

# Barcode clashing is done before datasets polishing in parallel, and then applied here
tib_all_PlusLif_v3_b1_v4_unique<-tib_all_PlusLif_v3_b1_v4%>%
  filter(barcode %in% all_ipcr_merged$barcode)

```

# Processing v4

```{r}
# offsets are computed after combining pDNA counts, from pDNA count columns, after filtering
auto_process_v4 <- function(tib_all, min_pdna, min_n_bc, avg_type) {

  # sum pdna counts across replicates
   tib_all    <- combine_cn(tib_all)

  # remove barcodes with pDNA reads < threshold
  tib_all    <- filter_cn(tib_all, min_pdna = min_pdna)


  # remove barcodes of elements with < threshold distinct barcodes
  tib_all    <- filter_barcodes(tib_all, min_n_bc = min_n_bc,  main = 'AllBC')

  # compute offsets
  offsets_all <- tib_all %>%
    dplyr::select(contains('dna_counts')) %>%
    colSums() / 1e6
  names(offsets_all) <- names(offsets_all) %>%
    str_replace('_counts', '')
  
  message('norm factors', offsets_all)
  
    
  # normalize to library sizes
  tib_all    <- normalize_counts(tib_all, offsets_all)

  # compute activities and summarize barcodes
  tib_all    <- compute_activities(tib_all)

  # summarize activities (weighted average)
  tib_all    <- summarize_barcodes(tib_all, type = avg_type)

#  if(missing(signal_t)) {

    # model combinatorial activity distributions
 #   signal_t     <- learn_min_activity(tib_comb)

 # }

 # # compute cooperativity
#  tib_comb_all <- compute_cooperativity(tib_comb, tib_basal, signal_t = signal_t, remove_inactive = FALSE)
 # tib_comb     <- compute_cooperativity(tib_comb, tib_basal, signal_t = signal_t, remove_inactive = TRUE)

  # add metadata
 # tib_basal    <- add_metadata(tib_basal, tib_dat = tib_dat, type = 'basal')
#  tib_comb     <- add_metadata(tib_comb, tib_dat = tib_dat, type = 'comb')
 # tib_comb_all <- add_metadata(tib_comb_all, tib_dat = tib_dat, type = 'comb')

  # return list of tibs
  return(tib_all)

}
```


```{r set params}
# min number of barcodes
min_n_bc <- 5
# min copy number
min_pdna <- 8


# load genomic coordinates of DAT library
tib_frag_coord <- read_dat_design('/home/f.comoglio/gitlab/epmatrix/data/tad_deconstr_design/designs/fc180419_dat_tfcp2l1.bed')
```

```{r autoprocess tibs}

# change from previous versions, now we process first barcode aggregation and activities computations, in a later step we calculate boosting.
tibs_tfcp2l1_v3_b1_v4        <- auto_process_v4( tib_all_PlusLif_v3_b1_v4_unique, 
                               min_pdna  = min_pdna, 
                               min_n_bc  = min_n_bc, 
                               avg_type  = 'avg')


```


## Postprocessing
### Separate libraries
```{r}
#split id from library identity
tibs_tfcp2l1_v3_b1_v4_idsplit <- tibs_tfcp2l1_v3_b1_v4 %>% 
  mutate(library=sapply(str_split(id, "_"), function(x) x[2])) %>% 
  mutate(id= sapply(str_split(id, "_"), function(x) x[1]))

## separate librariesib_basal_merged_idlib<-tib_basal_merged %>% 

tib_basal_tfcp2l1_v3_b1_v4 <- tibs_tfcp2l1_v3_b1_v4_idsplit %>% filter(library == "basal")

tib_comb_tfcp2l1_v3_b1_v4 <- tibs_tfcp2l1_v3_b1_v4_idsplit %>% filter(library == "comb")


```

### Compute Coop V4 standard

### Compute Coop V4 Empty
```{r}
compute_cooperativity_v4 <- function(tib_comb, tib_basal) {

 
  # generate key pair to extract basal activities of frag1/frag2
  tib_basal_act <- tib_comb %>%
    # add temp ids matching basal library id for each element of pair
    mutate(id_frag1 = paste0(frag1, strand1, frag1, strand1),
           id_frag2 = paste0(frag2, strand2, frag2, strand2)) %>%
    dplyr::select(id_frag1, id_frag2) %>%
    # extract activities basal frag 1
    left_join(dplyr::select(tib_basal, id, starts_with('activity')),
              by = c("id_frag1" = "id")) %>%
    # extract activities basal frag 2
    left_join(dplyr::select(tib_basal, id, starts_with('activity')),
              by = c("id_frag2" = "id"),
              suffix = c('_frag1', '_frag2')) %>%
    dplyr::select(-c(id_frag1, id_frag2))

  # add basal activities to comb
  tib_activities <- bind_cols(tib_comb, tib_basal_act)%>%
    mutate(boost=log2(activity_all/activity_all_frag2))%>%
    #mutate(coop_all=log2(activity_all/(activity_all_frag2+activity_all_frag1)))%>%
    dplyr::select(starts_with('activity'), boost)

  tib_comb <- bind_cols(tib_comb, tib_activities%>%select(-ends_with(c("br1","br2","br3","all"))))
  tib_comb
}

```


```{r}
tib_comb_tfcp2l_processed_v4B1<-compute_cooperativity_v4(tib_comb_tfcp2l1_v3_b1_v4,tib_basal_tfcp2l1_v3_b1_v4)

```

```{r}
tib_comb_tfcp2l_processed_v4B1%>% filter(frag1=="eempty")
```

```{r}
tib_comb_tfcp2l_processed_v4B1%>% 
  ggplot(aes(log2(activity_all_frag2),log2(activity_all)))+
          # geom_dotplot(binaxis = "y", stackdir = "center", binwidth = 0.03, size=0.1, width = 0.2)+
  geom_point(alpha=0.5)+
    theme_bw()+
  ggtitle("B3 V4 basal vs comb activity Tfcp2l1")+
  geom_abline(slope = 1, colour="red")
```


## Add metadata
```{r}
#function
add_metadata <- function(tib, tib_dat) {
tib <- tib %>%
      mutate(id1   = paste0(frag1, strand1),
             id2   = paste0(frag2, strand2),
             id    = paste0(id1, id2),
             class = paste0(substr(id1, 1, 1), substr(id2, 1, 1))) %>%
      # add genomic coordinates
      left_join(tib_dat, by = c('frag1' = 'frag')) %>%
      dplyr::rename(seqnames1 = seqnames, start1 = start, end1 = end) %>%
      left_join(tib_dat, by = c('frag2' = 'frag')) %>%
      dplyr::rename(seqnames2 = seqnames, start2 = start, end2 = end) %>%
      # group, trick essentially rowwise()
      group_by(id1, id2) %>%
      # distances for trans pairs is set to Inf
      mutate(dist = ifelse(seqnames1 == seqnames2,
                           min(abs(end1 - start2), abs(end2 - start1)), Inf)) %>%
      # reorder to final shape
      dplyr::select(id, class,
                    id1, frag1, strand1, seqnames1, start1, end1,
                    id2, frag2, strand2, seqnames2, start2, end2,
                    dist, starts_with('activ'), starts_with('boost'), starts_with('coop')) %>%
      ungroup

tib}
```

```{r}
#add metadata to each of the datasets
tib_comb_tfcp2l_processed_v4B1<-add_metadata(tib_comb_tfcp2l_processed_v4B1, tib_frag_coord)
```

```{r}
tib_comb_tfcp2l_processed_v4B1%>% 
  ggplot(aes(class,boost))+
          # geom_dotplot(binaxis = "y", stackdir = "center", binwidth = 0.03, size=0.1, width = 0.2)+
  geom_violin(alpha=0.5, aes(fill=class))+
  geom_boxplot(width=0.3)+
    theme_bw()+
  ggtitle("B3 V4 basal vs comb activity Tfcp2l1")

```

# Save dataset

```{r save tibbles, eval = TRUE}
save(list = c('tib_comb_tfcp2l_processed_v4B1', 'tib_comb_tfcp2l_processed_v4B1_ML24','tib_comb_tfcp2l_processed_v4B1_ML48'),
     file = '~/mydata/GitHub/EPCombinations/data/MMA20201010_epsure_tibs_tfcp2l1_V3_V4.RData')
```




### SessionInfo

```{r}
devtools::session_info()
```