---
title: "MMA20200727_NC_EPSuRE_Raw_to_Boost_V3pipeline_V5Norm"
author: "MMA"
date: "27/7/2020"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo       = TRUE, 
                      message    = FALSE, 
                      warning    = FALSE, 
                      cache      = TRUE,
                      autodep    = TRUE,
                      fig.align  = 'center', 
                      fig.width  = 6, 
                      fig.height = 6)
```

```{r libs and fun}
library(tidyverse)
library(here)
library(magrittr)
library(reshape2)
library(knitr)
library(gridExtra)
library(pander)
library(cowplot)
library(circlize)
library(datteRo)
library(GGally)
library(ggplot2)
library(ggpubr)

```

# Intro

In this document I am processing and merging together the iPCR data for EP libraries 10-13 (Klf2 and Nanog Comb and Basal, seq data E43) and 16-17 (NCs+Klf2, seq data E86) with the ecn data generated in E87 wich includes all the 6 libraries mentioned.

# Preprocessing
## iPCR 

We loaded basal and combinatorial iPCR libraries from the EP pipeline v3 (include empty fragments).

```{r process ipcr basal lib}


# load basal libraries
tib_basal_klf2  <- read_ipcr('/DATA/projects/epmatrix/mouse/DAT_EP010_EP013/ipcr/Results_v3/ipcr_E43_EP013_dat_klf2_basal_miseq_br1_unique_barcodes_frag_id_merged.tsv')
tib_basal_nanog <- read_ipcr('/DATA/projects/epmatrix/mouse/DAT_EP010_EP013/ipcr/Results_v3/ipcr_E43_EP011_dat_nanog_basal_miseq_br1_unique_barcodes_frag_id_merged.tsv')
tib_basal_NCklf2 <- read_ipcr('~/epmatrix/mouse/DAT_EP016_EP017/ipcr/Results_v3/E85_iPCR_EP016_BasalNC_miseq_br1_unique_barcodes_frag_id_merged.tsv')

allBasal_ipcr_merged<-bind_rows(tib_basal_klf2,tib_basal_nanog,tib_basal_NCklf2)

#count empty fragments
tib_basal_klf2%>% filter(frag1=="eempty")%>% nrow()
tib_basal_nanog%>% filter(frag1=="eempty")%>% nrow()
tib_basal_NCklf2%>% filter(frag1=="eempty")%>% nrow()


# count heterotypic

tib_basal_klf2%>% filter(frag1!=frag2 | strand1!=strand2)%>% nrow()
tib_basal_nanog%>% filter(frag1!=frag2 | strand1!=strand2)%>% nrow()
tib_basal_NCklf2%>% filter(frag1!=frag2 | strand1!=strand2)%>% nrow()

# remove heterotypic
tib_basal_klf2 <- remove_heterotypic(tib_basal_klf2)
tib_basal_nanog <- remove_heterotypic(tib_basal_nanog)
tib_basal_NCklf2 <- remove_heterotypic(tib_basal_NCklf2)

# fragment representation
count_fragments(tib_basal_klf2)
count_elements(tib_basal_klf2)
count_bc_reads(tib_basal_klf2)

count_fragments(tib_basal_nanog)
count_elements(tib_basal_nanog)
count_bc_reads(tib_basal_nanog)

count_fragments(tib_basal_NCklf2)
count_elements(tib_basal_NCklf2)
count_bc_reads(tib_basal_NCklf2)
```

```{r process ipcr comb lib}
# load combinatorial library
tib_comb_klf2  <- read_ipcr('/DATA/projects/epmatrix/mouse/DAT_EP010_EP013/ipcr/Results_v3/ipcr_E43_EP012_dat_klf2_combinatorial_miseq_br1_unique_barcodes_frag_id_merged.tsv')
tib_comb_nanog <- read_ipcr('/DATA/projects/epmatrix/mouse/DAT_EP010_EP013/ipcr/Results_v3/ipcr_E43_EP010_dat_nanog_combinatorial_miseq_br1_unique_barcodes_frag_id_merged.tsv')
tib_comb_NCKlf2 <- read_ipcr('~/epmatrix/mouse/DAT_EP016_EP017/ipcr/Results_v3/E85_iPCR_EP017_CombNC_miseq_br1_unique_barcodes_frag_id_merged.tsv')

allComb_ipcr_merged<-bind_rows(tib_comb_klf2,tib_comb_nanog,tib_comb_NCKlf2)

#count empty fragments
tib_comb_klf2%>% filter(frag1=="eempty")%>% nrow()
tib_comb_nanog%>% filter(frag1=="eempty")%>% nrow()
tib_comb_NCKlf2%>% filter(frag1=="eempty")%>% nrow()

# count homotypic
tib_comb_klf2%>% filter(frag1==frag2 & strand1==strand2)%>% nrow()
tib_comb_nanog%>% filter(frag1==frag2 & strand1==strand2)%>% nrow()
tib_comb_NCKlf2%>% filter(frag1==frag2 & strand1==strand2)%>% nrow()

#put appart empty fragments
tib_comb_klf2_empty<-tib_comb_klf2%>% filter(frag1=="eempty")
tib_comb_nanog_empty<-tib_comb_nanog%>% filter(frag1=="eempty")
tib_comb_NCKlf2_empty<-tib_comb_NCKlf2%>% filter(frag1=="eempty")

# mark and remove basal fragments from comb library. NOTE: It also removes empty fragments, that is why we put them appart before and bring them back afterwards.
tib_comb_klf2_het  <- mark_basal(tib_comb_klf2) %>%
  filter(!with_basal)
tib_comb_nanog_het  <- mark_basal(tib_comb_nanog) %>%
  filter(!with_basal)
tib_comb_NCKlf2_het<- mark_basal(tib_comb_NCKlf2) %>%
  filter(!with_basal)

# bring back empty elements
tib_comb_klf2_hetempty<-bind_rows(tib_comb_klf2_het,tib_comb_klf2_empty)
tib_comb_nanog_hetempty<-bind_rows(tib_comb_nanog_het,tib_comb_nanog_empty)
tib_comb_NCKlf2_hetempty<-bind_rows(tib_comb_NCKlf2_het,tib_comb_NCKlf2_empty)
```
### get list of unique barcodes
```{r}
#merge all to check proper barcode clashing
all_ipcr_merged<-bind_rows(allBasal_ipcr_merged,allComb_ipcr_merged)
# get unique barcodes before filtering
all_ipcr_merged_unique<-all_ipcr_merged %>%
  ungroup()%>%
  nest(-barcode ) %>% 
  mutate(n = map_dbl(data, nrow))%>%
  filter(n==1)%>%
  unnest()%>% 
  select(-n)
```

### Adding pDNA and cDNA data

For each iPCR library, we combined pDNA and cDNA data, retaining all barcodes with pDNA counts $\geq$ 1 and detected across all replicates.

```{r batch1 merge ipcr and ecn data}

###### To do Change all names and data loading

# read expression data (cdna)
tib_expr_br1 <- read_ecn('/DATA/projects/epmatrix/mouse/DAT_EP016_EP017/ecn/Results_v3/E87_NC_Klf2Nanog_EPSuRE_cDNA_BR1_barcode_counts.tsv', col_lab = 'cdna_counts_br1')
tib_expr_br2 <- read_ecn('/DATA/projects/epmatrix/mouse/DAT_EP016_EP017/ecn/Results_v3/E87_NC_Klf2Nanog_EPSuRE_cDNA_BR2_barcode_counts.tsv', col_lab = 'cdna_counts_br2')
tib_expr_br3 <- read_ecn('/DATA/projects/epmatrix/mouse/DAT_EP016_EP017/ecn/Results_v3/E87_NC_Klf2Nanog_EPSuRE_cDNA_BR3_barcode_counts.tsv', col_lab = 'cdna_counts_br3')

# read copy number data (pdna)
tib_cn_br1   <- read_ecn('/DATA/projects/epmatrix/mouse/DAT_EP016_EP017/ecn/Results_v3/E87_NC_Klf2Nanog_EPSuRE_pDNA_BR1_barcode_counts.tsv', col_lab = 'pdna_counts_br1')
tib_cn_br2  <- read_ecn('/DATA/projects/epmatrix/mouse/DAT_EP016_EP017/ecn/Results_v3/E87_NC_Klf2Nanog_EPSuRE_pDNA_BR2_barcode_counts.tsv', col_lab = 'pdna_counts_br2')

# merge datasets Klf2
tib_basal_merged_klf2_v3_b3  <- merge_ipcr_ecn(tib           = tib_basal_klf2, 
                                           cn_tib_list    = list(tib_cn_br1, tib_cn_br2),
                                           expr_tib_list  = list(tib_expr_br1, tib_expr_br2, tib_expr_br3))

tib_comb_merged_klf2_v3_b3   <- merge_ipcr_ecn(tib           = tib_comb_klf2_hetempty, 
                                           cn_tib_list    = list(tib_cn_br1, tib_cn_br2),
                                           expr_tib_list  = list(tib_expr_br1, tib_expr_br2, tib_expr_br3))

# merge datasets Klf2
tib_basal_merged_nanog_v3_b3 <- merge_ipcr_ecn(tib           = tib_basal_nanog, 
                                           cn_tib_list    = list(tib_cn_br1, tib_cn_br2),
                                           expr_tib_list  = list(tib_expr_br1, tib_expr_br2, tib_expr_br3))

tib_comb_merged_nanog_v3_b3  <- merge_ipcr_ecn(tib           = tib_comb_nanog_hetempty, 
                                           cn_tib_list    = list(tib_cn_br1, tib_cn_br2),
                                           expr_tib_list  = list(tib_expr_br1, tib_expr_br2, tib_expr_br3))

# merge datasets NCs
tib_basal_merged_NCklf2_v3_b3  <- merge_ipcr_ecn(tib           = tib_basal_NCklf2, 
                                           cn_tib_list    = list(tib_cn_br1, tib_cn_br2),
                                           expr_tib_list  = list(tib_expr_br1, tib_expr_br2, tib_expr_br3))

tib_comb_merged_NCklf2_v3_b3   <- merge_ipcr_ecn(tib           = tib_comb_NCKlf2_hetempty, 
                                           cn_tib_list    = list(tib_cn_br1, tib_cn_br2),
                                           expr_tib_list  = list(tib_expr_br1, tib_expr_br2, tib_expr_br3))
```

### Merging klf2 and nanog datasets b2
```{r}
tib_basal_merged_klf2_v3_b3_idlib<-tib_basal_merged_klf2_v3_b3 %>% 
  mutate(id=paste0(id, "_klf2basal"))

tib_comb_merged_klf2_v3_b3_idlib<-tib_comb_merged_klf2_v3_b3 %>% 
  mutate(id=paste0(id, "_klf2comb"))

tib_basal_merged_nanog_v3_b3_idlib<-tib_basal_merged_nanog_v3_b3 %>% 
  mutate(id=paste0(id, "_nanogbasal"))

tib_comb_merged_nanog_v3_b3_idlib<-tib_comb_merged_nanog_v3_b3 %>% 
  mutate(id=paste0(id, "_nanogcomb"))

tib_basal_merged_NCklf2_v3_b3_idlib<-tib_basal_merged_NCklf2_v3_b3 %>% 
  mutate(id=paste0(id, "_NCklf2basal"))

tib_comb_merged_NCklf2_v3_b3_idlib<-tib_comb_merged_NCklf2_v3_b3 %>% 
  mutate(id=paste0(id, "_NCklf2comb"))

# v3_b3_v4 stands for v3 EP-sure pipeline, batch 3 expression data generation (with NC libs), v4 data normalisation.

tib_all_NCklf2nanog_v3_b3_v4<-bind_rows(tib_basal_merged_klf2_v3_b3_idlib,tib_comb_merged_klf2_v3_b3_idlib,tib_basal_merged_nanog_v3_b3_idlib,tib_comb_merged_nanog_v3_b3_idlib,tib_basal_merged_NCklf2_v3_b3_idlib, tib_comb_merged_NCklf2_v3_b3_idlib )

#new barcode clashing removal
# This step is now crucial because there is no barcode clashing removal in the pipeline
#tib_all_NCklf2nanog_v3_b3_v4_unique <- tib_all_NCklf2nanog_v3_b3_v4 %>% ungroup()%>%
#  nest(-barcode ) %>% 
#  mutate(n = map_dbl(data, nrow))%>%
#  filter(n==1)%>%
#  unnest()%>% 
#  select(-n)

# Barcode clashing is done before datasets polishing in parallel, and then applied here
tib_all_NCklf2nanog_v3_b3_v4_unique<-tib_all_NCklf2nanog_v3_b3_v4%>%
  filter(barcode %in% all_ipcr_merged_unique$barcode)
```

# Barcode counts distributions

```{r pDNA_BarcodeCountDistributions}
tib_all_NCklf2nanog_v3_b3_v4_unique%>%
  ggplot(aes(pdna_counts_br1))+
  geom_histogram()

tib_all_NCklf2nanog_v3_b3_v4_unique%>%
  ggplot(aes(log10(pdna_counts_br1)))+
  geom_histogram()

tib_all_NCklf2nanog_v3_b3_v4_unique%>%
  ggplot(aes(pdna_counts_br2))+
  geom_histogram()


tib_all_NCklf2nanog_v3_b3_v4_unique%>%
  ggplot(aes(log10(pdna_counts_br2)))+
  geom_histogram()
```


```{r pDNA_BarcodeCountDistributions}
tib_all_NCklf2nanog_v3_b3_v4_unique%>%
  ggplot(aes(cdna_counts_br1))+
  ylim(c(0,4e+06))+
  geom_histogram()

tib_all_NCklf2nanog_v3_b3_v4_unique%>%
  mutate(cdna_counts_br1=cdna_counts_br1+1)%>%
  ggplot(aes(log10(cdna_counts_br1)))+
   ylim(c(0,2e+06))+
  geom_histogram(binwidth = 0.1)


tib_all_NCklf2nanog_v3_b3_v4_unique%>%
  ggplot(aes((cdna_counts_br2)))+
    ylim(c(0,4e+05))+
  geom_histogram()

tib_all_NCklf2nanog_v3_b3_v4_unique%>%
   mutate(cdna_counts_br2=cdna_counts_br2+1)%>%
  ggplot(aes(log10(cdna_counts_br2)))+
     ylim(c(0,2e+06))+
  geom_histogram(binwidth = 0.1)


tib_all_NCklf2nanog_v3_b3_v4_unique%>%
  ggplot(aes((cdna_counts_br3)))+
    ylim(c(0,4e+05))+
  geom_histogram()

tib_all_NCklf2nanog_v3_b3_v4_unique%>%
  mutate(cdna_counts_br3=cdna_counts_br3+1)%>%
  ggplot(aes(log10(cdna_counts_br3)))+
     ylim(c(0,2e+06))+
  geom_histogram(binwidth = 0.1)
```

```{r}
tib_all_NCklf2nanog_v3_b3_v4_unique%>%
      arrange(desc(pdna_counts_br1)) %>%
      mutate(n_cum = cumsum(pdna_counts_br1)) %>%
      ggplot(aes( 1 : length(n_cum), n_cum,)) +
        geom_line() +
        labs(x = 'Cumulative read counts', y = 'Cumulative # of barcodes') +
        theme_bw()

tib_all_NCklf2nanog_v3_b3_v4_unique%>%
      arrange(desc(pdna_counts_br2)) %>%
      mutate(n_cum = cumsum(pdna_counts_br2)) %>%
      ggplot(aes( 1 : length(n_cum), n_cum)) +
        geom_line() +
        labs(x = 'Cumulative read counts', y = 'Cumulative # of barcodes') +
        theme_bw()
```


```{r}
tib_all_NCklf2nanog_v3_b3_v4_unique%>%
  filter(!cdna_counts_br1==0)%>%
      arrange(desc(cdna_counts_br1)) %>%
      mutate(n_cum = cumsum(cdna_counts_br1)) %>%
      ggplot(aes( 1 : length(n_cum), n_cum,)) +
        geom_line() +
        labs(x = 'Cumulative read counts', y = 'Cumulative # of barcodes') +
        theme_bw()

tib_all_NCklf2nanog_v3_b3_v4_unique%>%
    filter(!cdna_counts_br2==0)%>%
      arrange(desc(cdna_counts_br2)) %>%
      mutate(n_cum = cumsum(cdna_counts_br2)) %>%
      ggplot(aes( 1 : length(n_cum), n_cum)) +
        geom_line() +
        labs(x = 'Cumulative read counts', y = 'Cumulative # of barcodes') +
        theme_bw()

tib_all_NCklf2nanog_v3_b3_v4_unique%>%
    filter(!cdna_counts_br3==0)%>%
      arrange(desc(cdna_counts_br3)) %>%
      mutate(n_cum = cumsum(cdna_counts_br3)) %>%
      ggplot(aes( 1 : length(n_cum), n_cum)) +
        geom_line() +
        labs(x = 'Cumulative read counts', y = 'Cumulative # of barcodes') +
        theme_bw()
```

# Processing v4

```{r}
# offsets are computed after combining pDNA counts, from pDNA count columns, after filtering
auto_process_v4 <- function(tib_all, min_pdna, min_n_bc, avg_type) {

  # sum pdna counts across replicates
   tib_all    <- combine_cn(tib_all)

  # remove barcodes with pDNA reads < threshold
  tib_all    <- filter_cn(tib_all, min_pdna = min_pdna)


  # remove barcodes of elements with < threshold distinct barcodes
  tib_all    <- filter_barcodes(tib_all, min_n_bc = min_n_bc,  main = 'AllBC')

  # compute offsets
  offsets_all <- tib_all %>%
    dplyr::select(contains('dna_counts')) %>%
    colSums() / 1e6
  names(offsets_all) <- names(offsets_all) %>%
    str_replace('_counts', '')
  
  message('norm factors', offsets_all)
  
    
  # normalize to library sizes
  tib_all    <- normalize_counts(tib_all, offsets_all)

  # compute activities and summarize barcodes
  tib_all    <- compute_activities(tib_all)

  # summarize activities (weighted average)
  tib_all    <- summarize_barcodes(tib_all, type = avg_type)

#  if(missing(signal_t)) {

    # model combinatorial activity distributions
 #   signal_t     <- learn_min_activity(tib_comb)

 # }

 # # compute cooperativity
#  tib_comb_all <- compute_cooperativity(tib_comb, tib_basal, signal_t = signal_t, remove_inactive = FALSE)
 # tib_comb     <- compute_cooperativity(tib_comb, tib_basal, signal_t = signal_t, remove_inactive = TRUE)

  # add metadata
 # tib_basal    <- add_metadata(tib_basal, tib_dat = tib_dat, type = 'basal')
#  tib_comb     <- add_metadata(tib_comb, tib_dat = tib_dat, type = 'comb')
 # tib_comb_all <- add_metadata(tib_comb_all, tib_dat = tib_dat, type = 'comb')

  # return list of tibs
  return(tib_all)

}
```

```{r}
# min number of barcodes
min_n_bc <- 5
# min copy number
min_pdna <- 8

tib_all_v3_b3_v4<-auto_process_v4(tib_all_NCklf2nanog_v3_b3_v4_unique,
                min_pdna  = min_pdna, 
                min_n_bc  = min_n_bc, 
                avg_type  = 'avg')
```


## Postprocessing
### Separate libraries
```{r}
tib_all_v3_b3_v4_idsplit <- tib_all_v3_b3_v4 %>% 
  mutate(library=sapply(str_split(id, "_"), function(x) x[2])) %>% 
  mutate(id= sapply(str_split(id, "_"), function(x) x[1]))


tib_basal_klf2_v3_b3_v4 <- tib_all_v3_b3_v4_idsplit %>% filter(library == "klf2basal")

tib_comb_klf2_v3_b3_v4 <- tib_all_v3_b3_v4_idsplit %>% filter(library == "klf2comb")

tib_basal_nanog_v3_b3_v4 <- tib_all_v3_b3_v4_idsplit %>% filter(library == "nanogbasal")

tib_comb_nanog_v3_b3_v4 <- tib_all_v3_b3_v4_idsplit %>% filter(library == "nanogcomb")

tib_basal_NCklf2_v3_b3_v4 <- tib_all_v3_b3_v4_idsplit %>% filter(library == "NCklf2basal")

tib_comb_NCklf2_v3_b3_v4 <- tib_all_v3_b3_v4_idsplit %>% filter(library == "NCklf2comb")


```



### Compute Coop V4 standard

### Compute Coop V4 Empty
```{r}
compute_cooperativity_v4 <- function(tib_comb, tib_basal) {

 
  # generate key pair to extract basal activities of frag1/frag2
  tib_basal_act <- tib_comb %>%
    # add temp ids matching basal library id for each element of pair
    mutate(id_frag1 = paste0(frag1, strand1, frag1, strand1),
           id_frag2 = paste0(frag2, strand2, frag2, strand2)) %>%
    dplyr::select(id_frag1, id_frag2) %>%
    # extract activities basal frag 1
    left_join(dplyr::select(tib_basal, id, starts_with('activity')),
              by = c("id_frag1" = "id")) %>%
    # extract activities basal frag 2
    left_join(dplyr::select(tib_basal, id, starts_with('activity')),
              by = c("id_frag2" = "id"),
              suffix = c('_frag1', '_frag2')) %>%
    dplyr::select(-c(id_frag1, id_frag2))

  # add basal activities to comb
  tib_activities <- bind_cols(tib_comb, tib_basal_act)%>%
    mutate(boost_all=log2(activity_all/activity_all_frag2))%>%
    #mutate(coop_all=log2(activity_all/(activity_all_frag2+activity_all_frag1)))%>%
    dplyr::select(starts_with('activity'), boost_all)

  tib_comb <- bind_cols(tib_comb, tib_activities%>% select(-activity_br1,-activity_br2,-activity_br3, -activity_all))
  tib_comb
}

```


```{r}
tib_comb_klf2_processed_v4B3<-compute_cooperativity_v4(tib_comb_klf2_v3_b3_v4,tib_basal_klf2_v3_b3_v4)
tib_comb_nanog_processed_v4B3<-compute_cooperativity_v4(tib_comb_nanog_v3_b3_v4,tib_basal_nanog_v3_b3_v4)
tib_comb_NCklf2_processed_v4B3<-compute_cooperativity_v4(tib_comb_NCklf2_v3_b3_v4,tib_basal_NCklf2_v3_b3_v4)


tib_comb_klf2_processed_v4B3%>% 
  ggplot(aes(library,boost_all))+
          # geom_dotplot(binaxis = "y", stackdir = "center", binwidth = 0.03, size=0.1, width = 0.2)+
  geom_boxplot()+
    theme_bw()+
  ylim(c(-2.5,2.5))+
  geom_boxplot()+
  ggtitle("B3 V4 Boost Klf2")

tib_comb_NCklf2_processed_v4B3%>% 
    filter((!substring(frag1, 1, 1)=="P"))%>%
  ggplot(aes(library,boost_all))+
          # geom_dotplot(binaxis = "y", stackdir = "center", binwidth = 0.03, size=0.1, width = 0.2)+
  geom_boxplot()+
    theme_bw()+
  ylim(c(-2.5,2.5))+
  geom_boxplot()+
  ggtitle("B3 V4 Boost NCKlf2")
```
## Save tibs

```{r save tibbles}
save(list = c('tib_comb_klf2_processed_v4B3', 'tib_comb_nanog_processed_v4B3', 'tib_comb_NCklf2_processed_v4B3', 'tib_comb_klf2_v3_b3_v4','tib_basal_klf2_v3_b3_v4','tib_comb_nanog_v3_b3_v4','tib_basal_nanog_v3_b3_v4', 'tib_comb_NCklf2_v3_b3_v4', 'tib_basal_NCklf2_v3_b3_v4' ),
     file = '~/mydata/GitLab/epmatrix/data/MMA20200804_NCKlf2Nanog_EPSuRE_v3_B3_v4.RData')
```

# Matrix

```{r}
load('~/mydata/GitLab/epmatrix/data/MMA20200804_NCKlf2Nanog_EPSuRE_v3_B3_v4.RData')
```


```{r ploot matrix function, include=FALSE}

#Function to plot the Boosting index Matrix 
plot_coop_heatmap_simple <- function(tib, classs = 'all' , skip_n_lab = 1) {

  # validate args
  if(!classs %in% c('all', 'EE', 'EP', 'PE', 'PP', 'EN', 'PN', 'NP', 'NN'))
    stop('Undefined class type. Please use one of EE, EP, PE or PP, or leave default for all pairs')

  if(classs != 'all') {
    # note: tib %>% filter(class == get('class')) fails
    # see https://stackoverflow.com/questions/40169949/filter-dataframe-using-global-variable-with-the-same-name-as-column-name

    tib     <- tib %>% filter(class %in% class)
  }

  # set color palette
  pal <- c('navyblue', 'gray95', 'orangered')

  # static viz


      # compute coop index range
      y_min <- min(tib[['boost']], na.rm = TRUE)
      y_max <- max(tib[['boost']], na.rm = TRUE)

      # coop index heatmap (additive)
      p_main <- tib %>%
        ggplot(aes(x = id1, y = id2)) +
        geom_tile(fill = 'gray40') +
        #        geom_tile(color = 'gray20', alpha = 0, size = 0.5) + ## prev version with stroke
        #        geom_tile(fill = rgb(53/255, 143/255, 168/255, 0.1)) + ## prev version with color outside coop idx palette
        geom_raster(data = tib, aes(x = id1, y = id2, fill = boost))

    


    # marginal barplots
      
      tib_p_top <- tib %>%
      select(id1, activity_all_frag1) %>%
      distinct() %>%
      mutate(activity_orig = log2(activity_all_frag1),
             shift    = min(activity_orig[!is.infinite(activity_orig)], na.rm = TRUE),
             activity = activity_orig)
      
    p_top <- tib_p_top %>%
      ggplot(aes(id1, activity)) +
      geom_bar(stat = 'identity') +
      ylim(c(-5,5))
     # scale_y_continuous(breaks = c(min(tib_p_top[['activity']][!is.infinite(tib_p_top[['activity']])]),
     #                               max(tib_p_top[['activity']])),
      #                   labels = c(signif(min(tib_p_top[['activity_orig']][!is.infinite(tib_p_top[['activity_orig']])]), 2),
       #                             signif(max(tib_p_top[['activity_orig']]), 2)))

    tib_p_right <- tib %>%
      select(id2, activity_all_frag2) %>%
      distinct() %>%
      mutate(activity_orig = log2(activity_all_frag2),
             shift    = min(activity_orig[!is.infinite(activity_orig)], na.rm = TRUE),
             activity = activity_orig )

    p_right <- tib_p_right %>%
      ggplot(aes(id2, activity)) +
      geom_bar(stat = 'identity') +
      ylim(c(-5,5))+
      #scale_y_continuous(breaks = c(min(tib_p_right[['activity']][!is.infinite(tib_p_right[['activity']])]),
      #                              max(tib_p_right[['activity']])),
                        # labels = c(signif(min(tib_p_right[['activity_orig']][!is.infinite(tib_p_right[['activity_orig']])]), 2),
                                   # signif(max(tib_p_right[['activity_orig']]), 2))) +
      coord_flip()

    # cosmetics
    p_main <- p_main +
      scale_fill_gradientn(colors = pal,
                           values = rescale(c(y_min, 0, y_max))) +
      labs(x = 'Enhancers', y = 'Promoters', fill = 'Boosting Index', caption = paste0('(n = ', nrow(tib), ' pairs)')) +
      theme(axis.text.x      = element_text(angle = 90, color =  c('black', rep('transparent', skip_n_lab))),
            axis.text.y      = element_text(color =  c('black', rep('transparent', 1))),
            panel.grid       = element_blank(),
            panel.background = element_rect(fill = "white"),
            legend.position  = 'bottom',
            text             = element_text(size = 13),)

    p_top <- p_top +
      labs(x = NULL,  y = 'Basal Activity (log2)') +
      theme_bw() +
      theme(axis.line.x  = element_line(size = 0, colour = "white"),
            axis.ticks.x = element_line(size = 0),
            axis.text.x  = element_blank(),
            panel.border = element_blank(),
            panel.grid   = element_blank(),
            text         = element_text(size = 13))
    
    p_right <- p_right +
      labs(x = NULL,  y = 'Basal Activity (log2)') +
      theme_bw() +
      theme(axis.line.y  = element_line(size = 0, colour = "white"),
            axis.ticks.y = element_line(size = 0),
            axis.text.y  = element_blank(),
            panel.border = element_blank(),
            panel.grid   = element_blank(),
            text         = element_text(size = 13))

    p_empty <- ggplot(mtcars, aes(wt, mpg)) +
      theme(line             = element_blank(),
            text             = element_blank(),
            title            = element_blank(),
            panel.background = element_rect(fill = NA, size = 0), plot.background = element_rect(size = 0))
    egg::ggarrange(p_top,
                   p_empty,
                   p_main,
                   p_right,
                   nrow = 2,
                   ncol = 2,
                   widths = c(3, 1),
                   heights = c(1, 3))



}
```

```{r, fig.width=7, fig.height=7}
tib_comb_klf2_processed_v4B3%>%
  mutate(id2=paste0(frag2,strand2))%>%
    mutate(id1=paste0(frag1,strand1))%>%
  mutate(class=paste0(substr(frag1,1,1), substr(frag2,1,1)))%>%
  mutate(boost=log2(activity_all/activity_all_frag2))%>%
  filter(class=="EP")%>%
  ggplot(aes(class,boost))+
  geom_violin(aes(fill=class), binaxis = "y", stackdir = "center", binwidth = 0.02)+
  geom_boxplot(width=0.3)+
  #facet_wrap(~id2)+
  theme_bw(base_size = 15)+
   ylim(c(-2.5,5))+
  ggtitle("Boosting index distribution Upstream")
```

```{r, fig.height=10 , fig.width=20, include=FALSE}
#make plot
Klf2_matrix_b3_v4<-tib_comb_klf2_processed_v4B3%>%
  mutate(id2=paste0(frag2,strand2))%>%
    mutate(id1=paste0(frag1,strand1))%>%
  mutate(class=paste0(substr(frag1,1,1), substr(frag2,1,1)))%>%
  mutate(boost=log2(activity_all/activity_all_frag2))%>%
  filter(boost!=(-Inf))%>%
  filter(boost!=(+Inf))%>%
  filter(class=="EP")%>%
  filter(log2(activity_all_frag2)!=(-Inf))%>%
  filter(log2(activity_all_frag1)!=(-Inf))%>%
plot_coop_heatmap_simple( class = c('EP'), skip_n_lab = 0)

Klf2_matrix_b3_v4
```

```{r, fig.height=10 , fig.width=20, include=FALSE}
#make plot
Klf2_matrix_b3_v4EE<-tib_comb_klf2_processed_v4B3%>%
  mutate(id2=paste0(frag2,strand2))%>%
    mutate(id1=paste0(frag1,strand1))%>%
  mutate(class=paste0(substr(frag1,1,1), substr(frag2,1,1)))%>%
  mutate(boost=log2(activity_all/activity_all_frag2))%>%
  filter(boost!=(-Inf))%>%
  filter(boost!=(+Inf))%>%
  filter(class=="EE")%>%
  filter(log2(activity_all_frag2)!=(-Inf))%>%
  filter(log2(activity_all_frag1)!=(-Inf))%>%
plot_coop_heatmap_simple( class = c('EE'), skip_n_lab = 0)

Klf2_matrix_b3_v4EE
```

```{r, fig.height=10 , fig.width=20, include=FALSE}
#make plot
Klf2_matrix_b3_v4EEshort<-tib_comb_klf2_processed_v4B3%>%
  mutate(id2=paste0(frag2,strand2))%>%
    mutate(id1=paste0(frag1,strand1))%>%
  mutate(class=paste0(substr(frag1,1,1), substr(frag2,1,1)))%>%
  mutate(boost=log2(activity_all/activity_all_frag2))%>%
  filter(boost!=(-Inf))%>%
  filter(boost!=(+Inf))%>%
  filter(class=="EE")%>%
  filter(log2(activity_all_frag2)!=(-Inf))%>%
  filter(log2(activity_all_frag1)!=(-Inf))%>%
  filter(frag1%in% c("E005","E006", "E007", "E021", "E030", "E046","E048", "E057", "E058", "E061" ))%>%
  filter(frag2%in% c("E005","E006", "E007", "E021", "E030", "E046","E048", "E057", "E058", "E061" ))%>%
plot_coop_heatmap_simple( class = c('EE'), skip_n_lab = 0)

Klf2_matrix_b3_v4EEshort
```

```{r, fig.height=10 , fig.width=20, include=FALSE}
#save matrix
pdf(file = "/DATA/usr/m.martinez.ara/GitLab/epmatrix/rproj/figures/klf2_b3_v4_EPmatrix_MMA20200803.pdf",   # The directory you want to save the file in
    width = 30, # The width of the plot in inches
    height = 15) # The height of the plot in inches
 print(Klf2_matrix_b3_v4)

dev.off()
```

## Matrix Negtive controls

```{r fig.height=10, fig.height=10}
tib_comb_NCklf2_processed_v4NormB3Class<-tib_comb_NCklf2_processed_v4B3%>%
  mutate(class1= ifelse(substring(frag1, 1, 1)=="P", "P", ifelse(frag1=="eempty", "e", "N" )))%>%
  mutate(class2= ifelse(substring(frag2, 1, 1)=="P", "P", ifelse(frag2=="eempty", "e", "N" )))%>%
  mutate(class=paste0(class1,class2))%>%
  filter(class!="PP")

tib_comb_klf2_processed_v4NormB3Class<-tib_comb_klf2_processed_v4B3%>%
  mutate(class1= ifelse(substring(frag1, 1, 1)=="P", "P", ifelse(frag1=="eempty", "e","E")))%>%
  mutate(class2= ifelse(substring(frag2, 1, 1)=="P", "P", ifelse(frag2=="eempty", "e","E")))%>%
  mutate(class=paste0(class1,class2))

tib_allNCKlf2_v3_b3_v4<-bind_rows(tib_comb_klf2_processed_v4NormB3Class,tib_comb_NCklf2_processed_v4NormB3Class)


```

```{r, fig.height=15 , fig.width=30, include=FALSE}

#Preprocessdata

#make plot
NCKlf2_matrix_b3_v4<-tib_allNCKlf2_v3_b3_v4%>%
  mutate(id2=paste0(frag2,strand2))%>%
    mutate(id1=paste0(frag1,strand1))%>%
  #mutate(class=paste0(substr(frag1,1,1), substr(frag2,1,1)))%>%
  mutate(boost=log2(activity_all/activity_all_frag2))%>%
  filter(boost!=(-Inf))%>%
  filter(boost!=(+Inf))%>%
  filter(class%in%c('NP','PN','NN', 'PP'))%>%
  filter(log2(activity_all_frag2)!=(-Inf))%>%
  filter(log2(activity_all_frag1)!=(-Inf))%>%
plot_coop_heatmap_simple( class = c('NP','PN','NN'), skip_n_lab = 0)

NCKlf2_matrix_b3_v4
```
## Avtivity vs boost

```{r Boosting_vs_promoter_activity, fig.width = 15, fig.height = 15}
tib_allNCKlf2_v3_b3_v4 %>%
  mutate(id2=paste0(frag2,strand2))%>%
  mutate(id1=paste0(frag1,strand1))%>%
  #mutate(class=paste0(substr(frag1,1,1), substr(frag2,1,1)))%>%
  mutate(boost=log2(activity_all/activity_all_frag2))%>%
  filter(boost!=(-Inf))%>%
  filter(boost!=(+Inf))%>%
  #filter(class %in% c("EP"))%>%
  ggplot(aes(log2(activity_all_frag2), boost)) +
  scale_color_gradient2(low = 'navyblue', mid='gray95', high = 'orangered', name="Activity cE (log2)", guide= guide_colourbar(barwidth = 1, barheight = 10))+
  geom_point(size=1, alpha=0.7, aes(color=log2(activity_all_frag1))) +
  #geom_jitter( size=1, alpha=0.7, color="black",shape = 21, aes(fill=boost)) +
  xlab('Basal Activity Promoter (log2)') +
  ylab('Boosting Index') +
  ggtitle("Basal Activity Promoter vs Boosting v3 B3 v4") +
  theme_bw(base_size=15) + 
  geom_hline(yintercept = 0)+
  stat_cor(method = "pearson")+
  geom_smooth(method = lm)+
  facet_wrap(~class)+
  #labs(caption =  "cEP (n=10516)") +
  #stat_function(fun = (function(x)log2(2*2^(x))), colour = "black")+
  #stat_function(fun = (function(x)log2(2^(x)/2)), colour = "black")+
  coord_fixed(ratio = 1)
  
```

```{r Boosting_vs_promoter_activity, fig.width = 15, fig.height = 15}
tib_allNCKlf2_v3_b3_v4 %>%
  mutate(id2=paste0(frag2,strand2))%>%
  mutate(id1=paste0(frag1,strand1))%>%
  #mutate(class=paste0(substr(frag1,1,1), substr(frag2,1,1)))%>%
  mutate(boost=log2(activity_all/activity_all_frag2))%>%
  filter(boost!=(-Inf))%>%
  filter(boost!=(+Inf))%>%
  filter(class %in% c("EP", "PP", "NP","NN"))%>%
  ggplot(aes(log2(activity_all_frag2), boost)) +
  scale_color_gradient2(low = 'navyblue', mid='gray95', high = 'orangered', name="Activity cE (log2)", guide= guide_colourbar(barwidth = 1, barheight = 10))+
  geom_point(size=1, alpha=0.7, aes(color=log2(activity_all_frag1))) +
  #geom_jitter( size=1, alpha=0.7, color="black",shape = 21, aes(fill=boost)) +
  xlab('Basal Activity Promoter (log2)') +
  ylab('Boosting Index') +
  ggtitle("Basal Activity Promoter vs Boosting v3 B3 v4") +
  theme_bw(base_size=15) + 
  geom_hline(yintercept = 0)+
  geom_vline(xintercept = (-2))+
  stat_cor(method = "pearson")+
  geom_smooth(method = lm)+
  facet_wrap(~class)+
  #labs(caption =  "cEP (n=10516)") +
  #stat_function(fun = (function(x)log2(2*2^(x))), colour = "black")+
  #stat_function(fun = (function(x)log2(2^(x)/2)), colour = "black")+
  coord_fixed(ratio = 1)
  
```


# Explore normalisation

## Compare PPs present in Klf2 comb and NCKlf2 comb
```{r}
SharedPPs<-inner_join(tib_comb_NCklf2_processed_v4B3,tib_comb_klf2_processed_v4B3, by= "id")

```

```{r, fig.height=10, fig.width=10}
SharedPPs%>%
  ggplot(aes(boost_all.x, boost_all.y))+
  geom_jitter(aes(color=log2(activity_all_frag2.x)))+
  facet_wrap(~frag2.x)+
  geom_abline(slope=1)+
  ggtitle("PP shared combinations boost")
```

```{r, fig.height=10, fig.width=10}
SharedPPs%>%
  mutate(boostx= log2(activity_all.x/activity_all_frag2.x),
         boosty= log2(activity_all.y/activity_all_frag2.y))%>%
  ggplot(aes(boostx, boosty))+
  geom_jitter(aes(color=log2(activity_all_frag2.x)))+
  facet_wrap(~frag2.x)+
  geom_abline(slope=1)+
  ggtitle("PP shared combinations boost no norm")
```

```{r, fig.height=10, fig.width=10}
SharedPPs%>%
  ggplot(aes(log2(activity_all.x), log2(activity_all.y)))+
  geom_jitter(aes(color=log2(activity_all_frag2.x)))+
  facet_wrap(~frag2.x)+
  geom_abline(slope=1)+
  ggtitle("PP shared combinations Activities no norm")
```

```{r, fig.height=10, fig.width=10}
SharedPPs%>%
  ggplot(aes(log2(activity_all.x), log2(activity_all.y)))+
  geom_jitter(aes(color=log2(activity_all_frag2.x)))+
  #facet_wrap(~frag2.x)+
  geom_abline(slope=1)+
  geom_smooth(method = "lm")+
  stat_cor(method = "pearson")+
  ggtitle("PP shared combinations Activities, v4 no norm to empty")
```

```{r, fig.height=8, fig.width=8}
SharedPPs%>%
  ggplot(aes(log2(activity_all.x), log2(activity_all.y)))+
  geom_jitter()+
  #facet_wrap(~frag2.x)+
  geom_abline(slope=1)+
  geom_smooth(method = "lm")+
  theme_bw(base_size = 15)+
  stat_cor(method = "pearson")+
  xlab("Activity library 1(log2)")+
  ylab("Activity library 2 (log2)")+
  ggtitle("PP shared combinations Activities")
```







```{r, fig.height=8, fig.width=8}
SharedPPs%>%
  ggplot(aes(log2(activity_all_frag2.x), log2(activity_all_frag2.y)))+
  geom_jitter()+
  geom_abline(slope=1)+
  geom_smooth(method = "lm")+
 theme_bw(base_size = 15)+
  stat_cor(method = "pearson")+
  xlab("Activity basal library 1(log2)")+
  ylab("Activity basal library 2 (log2)")+
  ggtitle("Shared Ps basal Activities")
```

```{r, fig.height=6, fig.width=8}
SharedPPs%>%
  ggplot(aes(log2(activity_all_frag2.x), log2(activity_all_frag2.y)))+
  geom_jitter(aes(color=log2(activity_all_frag2.x)))+
  geom_abline(slope=1)+
  geom_smooth(method = "lm")+
  stat_cor(method = "pearson")+
  ggtitle("PP shared Basal Activities, v4 No norm to empty")
```

## Conclusion

V4 normalisation without normalising to empty fragments works very well, PPs shared between libraries show very similar activities and shared basal elements also look very similar. Normalising to empty fragments is probably just introducing noise in the data because in this case the empty fragment activities of all libraries are within noise range of each other. In B1 and B2 normalising to empty fragments seemed more important.

