---
title: "Figure4_EPcombinations"
author: "MMA"
date: "2/4/2020"
output:
  pdf_document:
    fig_caption: yes
    toc: yes
  html_document: 
    toc: yes
---

```{r setup, message=FALSE, warning=FALSE, include=FALSE}
knitr::opts_chunk$set(
	fig.align = "center",
	fig.path = "~/mydata/GitLab/epmatrix/rproj/FiguresPaper/Figure4/",
	fig.pos = "H",
	message = FALSE,
	warning = FALSE,
	dev = c("pdf"),
	dpi = 500,
	include = TRUE
)
```

#load libraries
```{r libs and fun, include=FALSE}
library(tidyverse)
library(magrittr)
library(reshape2)
library(knitr)
library(datteRo)
library(GGally)
library(ggplot2)
library(gridExtra)
library(GGally)
library(scales)
library(heatmaply)
library(here)
library(ggpubr)
library(ggbeeswarm)
library(onewaytests)
```


# Load data
```{r}
#load klf2 and NCKLF2  
load('~/mydata/GitHub/EPCombinations/data/MMA20201204_NormAssays_Clusters.RData')
#The 2 libs are contained in the same 
#save(list = c('tib_Split_NCS_avgSPLIT_V2Adj_all', 'allKlfcombNCAdj_all', 'Klf2ClustersKmeans4' ),
#     file = '~/mydata/GitLab/epmatrix/data/MMA20201204_NormAssays_Clusters.RData')

# Load Nanog tibs
load('~/mydata/GitHub/EPCombinations/data/MMA20210323_NormResc_Nanog.RData')

#load Tfcp2l1 tibs
load('~/mydata/GitHub/EPCombinations/data/MMA20210323_NormResc_Tfcp2l1.RData')
```


```{r}

tfcp2l1_comball<-allcomb_tfcp2Adj_all %>%
    #filter(library=="tfcp2l1comb")%>%
  mutate(DAT="Tfcp2l1")%>%
  mutate(idDAt=paste(id,"Tfcp2l1",sep = "_"))

Klf2_comball<-allKlfcombNCAdj_all %>%
  filter(library=="klf2comb")%>%
  mutate(DAT="klf2")%>%
  mutate(id1=paste0(frag1,strand1))%>%
  mutate(idDAt=paste(id,"Klf2",sep = "_"))

Nanog_comball<-allcomb_NanogAdj_all %>%
 filter(library=="nanogcomb")%>%
  mutate(DAT="Nanog")%>%
  mutate(idDAt=paste(id,"Nanog",sep = "_"))

all_comball<-bind_rows(tfcp2l1_comball,Klf2_comball,Nanog_comball)

```


# Correlation mean boost vs boost
## Both orientations
```{r General_effect_enhancers, fig.width = 5, fig.height = 5}
meanboostperE<-all_comball%>%
  filter(class=="EP")%>%
  filter(strand2=="+")%>%
  mutate(frag1=paste0(frag1,"_", DAT))%>%
  mutate(frag2=paste0(frag2,"_", DAT))%>%
  mutate(id2=paste0(id2,"_", DAT))%>%
  filter(!BoostPerPmed%in%c(NA,NaN,Inf,-Inf))%>%
  group_by(frag1)%>%
  summarise(meanboost=mean(BoostPerPmed),n=n())

all_comball%>%
  filter(class=="EP")%>%
  mutate(frag1=paste0(frag1,"_", DAT))%>%
  mutate(frag2=paste0(frag2,"_", DAT))%>%
  mutate(id2=paste0(id2,"_", DAT))%>%
  filter(!BoostPerPmed%in%c(NA,NaN,Inf,-Inf))%>%
  left_join(meanboostperE)%>%
  filter(n>5)%>%
  #filter(log2(activity_all)>-1)%>%
  ggplot(aes(meanboost,BoostPerPmed))+
  geom_point(size=0.5, alpha=0.5)+
  #stat_cor()+
  stat_cor()+
  xlab("meanboost Enhancer (log2)")+
  ggtitle("General effect candidate enhancer")+
  geom_smooth(method="lm")+
  theme_bw()+
  labs(caption = "EP (n=9418), E (n=304)")+
  ylab("Boost index (log2)")+
  xlab("General effect candidate Enhancer (log2)")
  


```
```{r General_effect_enhancers2, fig.width = 5, fig.height = 5}


all_comball%>%
  filter(class=="EP")%>%
  mutate(frag1=paste0(frag1,"_", DAT))%>%
  mutate(frag2=paste0(frag2,"_", DAT))%>%
  mutate(id2=paste0(id2,"_", DAT))%>%
  filter(!BoostPerPmed%in%c(NA,NaN,Inf,-Inf))%>%
  left_join(meanboostperE)%>%
  filter(n>5)%>%
  #filter(log2(activity_all)>-1)%>%
  filter(meanboost>1)%>%
  ggplot(aes(meanboost,BoostPerPmed))+
  geom_point(size=0.5, alpha=0.5)+
  #stat_cor()+
  stat_cor(aes(label = paste(..rr.label.., ..p.label.., sep = "~`,`~")))+
  xlab("meanboost Enhancer (log2)")+
  ggtitle("General Effect Highly Active Enhancers")+
  geom_smooth(method="lm")+
  theme_bw()+
  labs(caption = "EP (n=1,563), E (n=46)")+
  ylab("Boost index (log2)")+
  xlab("General effect Enhancer (log2)")
  


```

## AVG
```{r General_effect_enhancers_AVG, fig.width = 5, fig.height = 5}
meanboostperE<-all_comball%>%
  filter(class=="EP")%>%
  filter(strand2=="+")%>%
  mutate(frag1=paste0(frag1,"_", DAT))%>%
  mutate(frag2=paste0(frag2,"_", DAT))%>%
  mutate(id2=paste0(id2,"_", DAT))%>%
  filter(!BoostPerPmed%in%c(NA,NaN,Inf,-Inf))%>%
  group_by(frag1)%>%
  summarise(meanboost=mean(BoostPerPmed),n=n())

all_comball%>%
  filter(class=="EP")%>%
  filter(strand2=="+")%>%
  mutate(frag1=paste0(frag1,"_", DAT))%>%
  mutate(frag2=paste0(frag2,"_", DAT))%>%
  mutate(id2=paste0(id2,"_", DAT))%>%
  filter(!BoostPerPmed%in%c(NA,NaN,Inf,-Inf))%>%
  group_by(frag1, frag2)%>%
  summarise(BoostPerPmedAvgOrien=mean(BoostPerPmed))%>%
  left_join(meanboostperE)%>%
  filter(n>5)%>%
  #filter(log2(activity_all)>-1)%>%
  ggplot(aes(meanboost,BoostPerPmedAvgOrien))+
  geom_point(size=0.5, alpha=0.5)+
  #stat_cor()+
  stat_cor()+
  xlab("meanboost Enhancer (log2)")+
  ggtitle("General effect candidate enhancer")+
  geom_smooth(method="lm")+
  theme_bw()+
  labs(caption = "EP (n=3342), E (n=304)")+
  ylab("Boost index (log2)")+
  xlab("General effect candidate Enhancer (log2)")
  


```


## Assumption of equal variances within groups

```{r, fig.width=25, fig.height=5}
all_comball%>% 
  filter(class=="EP")%>%
  filter(strand2=="+")%>%
  #filter(log2(activity_all_frag2)>-1)%>%
  mutate(frag1=paste0(frag1,"_", DAT))%>%
  mutate(frag2=paste0(frag2,"_", DAT))%>%
  mutate(id1=paste0(id1,"_", DAT))%>%
     filter(!BoostPerPmed%in%c(NA,NaN,Inf,-Inf))%>%
   mutate(BoostPerPmed1=log2(activity_br1/Basal_median))%>%
    mutate(BoostPerPmed2=log2(activity_br2/Basal_median))%>%
    mutate(BoostPerPmed3=log2(activity_br3/Basal_median))%>%
  mutate(activity_all=log2(activity_all))%>%
  rowwise()%>%
  mutate(BiolVar=var(c(BoostPerPmed1,BoostPerPmed2, BoostPerPmed3)))%>%
  filter(DAT=="klf2")%>%
  ggplot(aes(id1, log10(BiolVar)))+
   geom_quasirandom()
  
```

```{r, fig.width=5, fig.height=5}
all_comball%>% 
  filter(class=="EP")%>%
  filter(strand2=="+")%>%
  #filter(log2(activity_all_frag2)>-1)%>%
  mutate(frag1=paste0(frag1,"_", DAT))%>%
  mutate(frag2=paste0(frag2,"_", DAT))%>%
  mutate(id1=paste0(id1,"_", DAT))%>%
     filter(!BoostPerPmed%in%c(NA,NaN,Inf,-Inf))%>%
   mutate(BoostPerPmed1=log2(activity_br1/Basal_median))%>%
    mutate(BoostPerPmed2=log2(activity_br2/Basal_median))%>%
    mutate(BoostPerPmed3=log2(activity_br3/Basal_median))%>%
  mutate(activity_all=log2(activity_all))%>%
  rowwise()%>%
  mutate(BiolVar=var(c(BoostPerPmed1,BoostPerPmed2, BoostPerPmed3)))%>%
  filter(id1%in%c("E097-_Klf2","E046+_Klf2","E070+_Klf2" ))%>%
  #filter(DAT=="klf2")%>%
  ggplot(aes(id1, log10(BiolVar)))+
   geom_quasirandom()
  
```
```{r, fig.width=25, fig.height=5}
all_comball%>% 
  filter(class=="EP")%>%
  filter(strand2=="+")%>%
  #filter(log2(activity_all_frag2)>-1)%>%
  mutate(frag1=paste0(frag1,"_", DAT))%>%
  mutate(frag2=paste0(frag2,"_", DAT))%>%
  mutate(id1=paste0(id1,"_", DAT))%>%
     filter(!BoostPerPmed%in%c(NA,NaN,Inf,-Inf))%>%
   mutate(BoostPerPmed1=log2(activity_br1/Basal_median))%>%
    mutate(BoostPerPmed2=log2(activity_br2/Basal_median))%>%
    mutate(BoostPerPmed3=log2(activity_br3/Basal_median))%>%
  mutate(activity_all=log2(activity_all))%>%
  rowwise()%>%
  mutate(BiolVar=var(c(BoostPerPmed1,BoostPerPmed2, BoostPerPmed3)))%>%
  filter(DAT=="klf2")%>%
  ggplot(aes(id2, log10(BiolVar)))+
   geom_quasirandom()
  
all_comball%>% 
  filter(class=="EP")%>%
  filter(strand2=="+")%>%
  #filter(log2(activity_all_frag2)>-1)%>%
  mutate(frag1=paste0(frag1,"_", DAT))%>%
  mutate(frag2=paste0(frag2,"_", DAT))%>%
  mutate(id1=paste0(id1,"_", DAT))%>%
     filter(!BoostPerPmed%in%c(NA,NaN,Inf,-Inf))%>%
   mutate(BoostPerPmed1=log2(activity_br1/Basal_median))%>%
    mutate(BoostPerPmed2=log2(activity_br2/Basal_median))%>%
    mutate(BoostPerPmed3=log2(activity_br3/Basal_median))%>%
  mutate(activity_all=log2(activity_all))%>%
  rowwise()%>%
  mutate(BiolVar=var(c(BoostPerPmed1,BoostPerPmed2, BoostPerPmed3)))%>%
  filter(DAT=="klf2")%>%
  filter(id2=="P003+")%>%
  ggplot(aes(id2, (BiolVar)))+
   geom_quasirandom()

```

We indeed see that the variances between groups are very wild in the enhancers. Thus we can't use the ANOVA test.

## normality of variances within groups 
```{r, fig.width=10, fig.height=5}
all_comball%>% 
  filter(class=="EP")%>%
  filter(strand2=="+")%>%
  #filter(log2(activity_all_frag2)>-1)%>%
  mutate(frag1=paste0(frag1,"_", DAT))%>%
  mutate(frag2=paste0(frag2,"_", DAT))%>%
  mutate(id1=paste0(id1,"_", DAT))%>%
     filter(!BoostPerPmed%in%c(NA,NaN,Inf,-Inf))%>%
   mutate(BoostPerPmed1=log2(activity_br1/Basal_median))%>%
    mutate(BoostPerPmed2=log2(activity_br2/Basal_median))%>%
    mutate(BoostPerPmed3=log2(activity_br3/Basal_median))%>%
  mutate(activity_all=log2(activity_all))%>%
  #rowwise()%>%
  #mutate(BiolVar=var(c(BoostPerPmed1,BoostPerPmed2, BoostPerPmed3)))%>%
  select(id1, frag2, id,BoostPerPmed1, BoostPerPmed2, BoostPerPmed3 )%>%
  pivot_longer(BoostPerPmed1:BoostPerPmed3, names_to = "rep", values_to = "BoostAll")%>%
  #filter(DAT=="klf2")%>%
  filter(id1=="E046+_Klf2")%>%
  ggplot(aes(frag2, BoostAll))+
   geom_quasirandom()

all_comball%>% 
  filter(class=="EP")%>%
  filter(strand2=="+")%>%
  #filter(log2(activity_all_frag2)>-1)%>%
  mutate(frag1=paste0(frag1,"_", DAT))%>%
  mutate(frag2=paste0(frag2,"_", DAT))%>%
  mutate(id1=paste0(id1,"_", DAT))%>%
     filter(!BoostPerPmed%in%c(NA,NaN,Inf,-Inf))%>%
   mutate(BoostPerPmed1=log2(activity_br1/Basal_median))%>%
    mutate(BoostPerPmed2=log2(activity_br2/Basal_median))%>%
    mutate(BoostPerPmed3=log2(activity_br3/Basal_median))%>%
  mutate(activity_all=log2(activity_all))%>%
  #rowwise()%>%
  #mutate(BiolVar=var(c(BoostPerPmed1,BoostPerPmed2, BoostPerPmed3)))%>%
  select(id1, frag2, id,BoostPerPmed1, BoostPerPmed2, BoostPerPmed3 )%>%
  pivot_longer(BoostPerPmed1:BoostPerPmed3, names_to = "rep", values_to = "BoostAll")%>%
  #filter(DAT=="klf2")%>%
  filter(id1=="E097-_Klf2")%>%
  ggplot(aes(frag2, BoostAll))+
   geom_quasirandom()

all_comball%>% 
  filter(class=="EP")%>%
  filter(strand2=="+")%>%
  #filter(log2(activity_all_frag2)>-1)%>%
  mutate(frag1=paste0(frag1,"_", DAT))%>%
  mutate(frag2=paste0(frag2,"_", DAT))%>%
  mutate(id1=paste0(id1,"_", DAT))%>%
     filter(!BoostPerPmed%in%c(NA,NaN,Inf,-Inf))%>%
   mutate(BoostPerPmed1=log2(activity_br1/Basal_median))%>%
    mutate(BoostPerPmed2=log2(activity_br2/Basal_median))%>%
    mutate(BoostPerPmed3=log2(activity_br3/Basal_median))%>%
  mutate(activity_all=log2(activity_all))%>%
  #rowwise()%>%
  #mutate(BiolVar=var(c(BoostPerPmed1,BoostPerPmed2, BoostPerPmed3)))%>%
  select(frag1,id1, frag2, id,BoostPerPmed1, BoostPerPmed2, BoostPerPmed3 )%>%
  pivot_longer(BoostPerPmed1:BoostPerPmed3, names_to = "rep", values_to = "BoostAll")%>%
  #filter(DAT=="klf2")%>%
  filter(frag1=="E070_Klf2")%>%
  ggplot(aes(frag2, BoostAll))+
   geom_quasirandom()
  
all_comball%>% 
  filter(class=="EP")%>%
  filter(strand2=="+")%>%
  #filter(log2(activity_all_frag2)>-1)%>%
  mutate(frag1=paste0(frag1,"_", DAT))%>%
  mutate(frag2=paste0(frag2,"_", DAT))%>%
  mutate(id1=paste0(id1,"_", DAT))%>%
     filter(!BoostPerPmed%in%c(NA,NaN,Inf,-Inf))%>%
   mutate(BoostPerPmed1=log2(activity_br1/Basal_median))%>%
    mutate(BoostPerPmed2=log2(activity_br2/Basal_median))%>%
    mutate(BoostPerPmed3=log2(activity_br3/Basal_median))%>%
  mutate(activity_all=log2(activity_all))%>%
  #rowwise()%>%
  #mutate(BiolVar=var(c(BoostPerPmed1,BoostPerPmed2, BoostPerPmed3)))%>%
  select(frag1,id1, frag2, id,BoostPerPmed1, BoostPerPmed2, BoostPerPmed3 )%>%
  pivot_longer(BoostPerPmed1:BoostPerPmed3, names_to = "rep", values_to = "BoostAll")%>%
  #filter(DAT=="klf2")%>%
  filter(frag1=="E030_Klf2")%>%
  ggplot(aes(frag2, BoostAll))+
   geom_quasirandom()
  
```



## Welch test Orientation AVG enhancers

Same but taking both orientations of enhancers.

```{r}
##Neccessary chunk moved around
VarPErE_AvgOr_within<-all_comball%>% 
  filter(class=="EP")%>%
  filter(strand2=="+")%>%
  #filter(log2(activity_all_frag2)>-1)%>%
  mutate(frag1=paste0(frag1,"_", DAT))%>%
  mutate(frag2=paste0(frag2,"_", DAT))%>%
  mutate(id1=paste0(id1,"_", DAT))%>%
     filter(!BoostPerPmed%in%c(NA,NaN,Inf,-Inf))%>%
   mutate(BoostPerPmed1=log2(activity_br1/Basal_median))%>%
    mutate(BoostPerPmed2=log2(activity_br2/Basal_median))%>%
    mutate(BoostPerPmed3=log2(activity_br3/Basal_median))%>%
  mutate(activity_all=log2(activity_all))%>%
  pivot_longer(c(BoostPerPmed1, BoostPerPmed2, BoostPerPmed3), names_to = "rep", values_to = "BoostperPmedall")%>%
  group_by(frag1, id2)%>%
  summarise(BiolVarAvgOr_within=var(BoostperPmedall), nrep=n())%>%
  filter(!BiolVarAvgOr_within%in%c(NA,NaN,Inf,-Inf))%>%
  ungroup()%>%
  group_by(frag1)%>%
  summarise(n2=n(),WithinVariance=sum(BiolVarAvgOr_within)/n2, k=sum(nrep))
  

VarPErE_AvgOr_between<-all_comball%>% 
  filter(class=="EP")%>%
  filter(strand2=="+")%>%
  #filter(log2(activity_all_frag2)>-1)%>%
  mutate(frag1=paste0(frag1,"_", DAT))%>%
  mutate(frag2=paste0(frag2,"_", DAT))%>%
    mutate(id2=paste0(id2,"_", DAT))%>%
  filter(!BoostPerPmed%in%c(NA,NaN,Inf,-Inf))%>%  
   mutate(BoostPerPmed1=log2(activity_br1/Basal_median))%>%
    mutate(BoostPerPmed2=log2(activity_br2/Basal_median))%>%
    mutate(BoostPerPmed3=log2(activity_br3/Basal_median))%>%
  mutate(activity_all=log2(activity_all))%>%
  group_by(frag1, id2)%>%
  summarise(BoostPerPmedAvgOr=mean(BoostPerPmed))%>%
  ungroup()%>%
  group_by(frag1)%>%
  summarise(variance=var(BoostPerPmedAvgOr), n=n())%>%
  left_join(VarPErE_AvgOr_within)%>%
  mutate(axis="y(enhancers)")%>%
  filter(!variance%in%c(NA,NaN,Inf,-Inf))%>%
  filter(!WithinVariance%in%c(NA,NaN,Inf,-Inf))%>%
  filter(n>5)%>%
  mutate(p_pvalue= pf(variance/WithinVariance, n-1, k-n, lower.tail = FALSE))%>%
  mutate(correctedPvalue=p.adjust(p_pvalue, "BH"))
```


```{r include=FALSE}
## Alternative F test to ANOVA Welch F test used when data is heterocedastic as it is our case.


forWelch<-all_comball%>% 
  filter(class=="EP")%>%
  filter(strand2=="+")%>%
  #filter(log2(activity_all_frag2)>-1)%>%
  mutate(frag1=paste0(frag1,"_", DAT))%>%
  mutate(frag2=paste0(frag2,"_", DAT))%>%
    mutate(id2=paste0(id2,"_", DAT))%>%
  filter(!BoostPerPmed%in%c(NA,NaN,Inf,-Inf))%>%  
   mutate(BoostPerPmed1=log2(activity_br1/Basal_median))%>%
    mutate(BoostPerPmed2=log2(activity_br2/Basal_median))%>%
    mutate(BoostPerPmed3=log2(activity_br3/Basal_median))%>%
  mutate(activity_all=log2(activity_all))%>%
  pivot_longer(c(BoostPerPmed1, BoostPerPmed2, BoostPerPmed3), names_to = "rep", values_to = "BoostperPmedall")%>%
  filter(frag1%in%VarPErE_AvgOr_between$frag1)


WelchTestEnhancers<-tibble(frag1=NA, pvalue=NA, fstatistic=NA, df1=NA, df2=NA )

for (i in unique(forWelch$frag1)) {
  wFtest<-forWelch%>%
    filter(frag1==i)%>%
    welch.test(BoostperPmedall ~ frag2, data = .)
  
  WelchTestEnhancers<-WelchTestEnhancers%>%
    add_row(frag1=i, pvalue=wFtest$p.value, fstatistic=wFtest$statistic, df1=wFtest$parameter[1], df2=wFtest$parameter[2])
}

### Both seem to work better than the anova as they pick up more significant cases than the Anova
```




```{r}
WelchTestEnhancers%>%
  ggplot(aes(pvalue))+
  geom_histogram()

```


```{r}
#WelchTestEnhancers%>%
#    mutate(CorrPvalue=p.adjust(pvalue, "BH"))%>%
#  mutate(DAT=sapply(str_split(frag1, "_"), function(x) x[2]))%>%
#  filter(DAT=="klf2")%>%
#  arrange(frag1)

WelchTestEnhancers%>%
  mutate(CorrPvalue=p.adjust(pvalue, "BH"))%>%
  mutate(Axis="enhancers")%>%
  ggplot(aes(Axis, log10(fstatistic)))+
  geom_quasirandom(size=0.5, varwidth = TRUE, dodge.width=0.4,aes(color=CorrPvalue<0.01))+
  theme_bw()
```
```{r Fstatistic_vs_MeanBoost, fig.height=5, fig.width=8}
#WelchTestEnhancers%>%
#    mutate(CorrPvalue=p.adjust(pvalue, "BH"))%>%
#  mutate(DAT=sapply(str_split(frag1, "_"), function(x) x[2]))%>%
#  filter(DAT=="klf2")%>%
#  arrange(frag1)
VarPErE_AvgOr_onlyVar<-all_comball%>% 
  ungroup()%>%
  filter(class=="EP")%>%
  filter(strand2=="+")%>%
  #filter(log2(activity_all_frag2)>-1)%>%
  mutate(frag1=paste0(frag1,"_", DAT))%>%
    mutate(id1=paste0(id1,"_", DAT))%>%
  mutate(frag2=paste0(frag2,"_", DAT))%>%
    mutate(id2=paste0(id2,"_", DAT))%>%
  filter(!BoostPerPmed%in%c(NA,NaN,Inf,-Inf))%>% 
  group_by(id2,frag1)%>%
  summarise(BoostPerPmed=mean(BoostPerPmed),activity_all_frag2=mean(activity_all_frag2),activity_all_frag1=mean(activity_all_frag1),class=unique(class))%>%
  group_by(frag1)%>%
  summarise(variance=var(BoostPerPmed), nsize=n(), meanboost=mean(BoostPerPmed))

#VarPErE_AvgOr_onlyVar<-VarPErE_AvgOr_between%>%dplyr::select(variance, frag1)

all_comball%>% 
ungroup()%>%
  filter(class=="EP")%>%
  filter(strand2=="+")%>%
  #filter(log2(activity_all_frag2)>-1)%>%
  mutate(frag1=paste0(frag1,"_", DAT))%>%
    mutate(id1=paste0(id1,"_", DAT))%>%
  mutate(frag2=paste0(frag2,"_", DAT))%>%
    mutate(id2=paste0(id2,"_", DAT))%>%
  filter(!BoostPerPmed%in%c(NA,NaN,Inf,-Inf))%>% 
    filter(frag1=="E030_klf2")

WelchTestEnhancers%>%
  mutate(CorrPvalue=p.adjust(pvalue, "BH"))%>%
  mutate(Axis="enhancers")%>%
  left_join(VarPErE_AvgOr_onlyVar, by=c("frag1"))%>%
  ggplot(aes(meanboost, log10(fstatistic), color=CorrPvalue<0.05))+
  #geom_quasirandom(size=0.5, varwidth = TRUE, dodge.width=0.4,aes(color=CorrPvalue<0.05))+
  geom_point(aes(size=nsize), alpha=0.5)+
  theme_bw()+
  labs(caption = "233 cRE (combinations>5)")
```
```{r Variance_vs_MeanBoost, fig.height=5, fig.width=8}
WelchTestEnhancers%>%
  mutate(CorrPvalue=p.adjust(pvalue, "BH"))%>%
  mutate(Axis="enhancers")%>%
  left_join(VarPErE_AvgOr_onlyVar, by=c("frag1"))%>%
  ggplot(aes(meanboost, log10(variance), color=CorrPvalue<0.05))+
  #geom_quasirandom(size=0.5, varwidth = TRUE, dodge.width=0.4,aes(color=CorrPvalue<0.05))+
  geom_point(aes(size=nsize), alpha=0.5)+
  theme_bw()

```
```{r }
WelchTestEnhancers%>%
  mutate(CorrPvalue=p.adjust(pvalue, "BH"))%>%
  mutate(significant=ifelse(CorrPvalue<0.05,1,0))%>%
  filter(!is.na(significant))%>%
  summarise(n=n(), PercSig=sum(significant)/n, nsignif=sum(significant))
```


```{r BarplotVariancepvalue1, fig.height=10, fig.width=5}
WelchTestEnhancers%>%
  mutate(CorrPvalue=p.adjust(pvalue, "BH"))%>%
  mutate(Axis="enhancers")%>%
  left_join(VarPErE_AvgOr_onlyVar, by=c("frag1"))%>%
  mutate(DAT=sapply(str_split(frag1, "_"), function(x) x[2]))%>%
  mutate(frag1=sapply(str_split(frag1, "_"), function(x) x[1]))%>% 
  filter(DAT=="klf2")%>%
  ggplot(aes(reorder(frag1, meanboost),meanboost))+
  #geom_quasirandom(size=0.5, varwidth = TRUE, dodge.width=0.4,aes(color=CorrPvalue<0.05))+
  geom_bar(stat="identity")+
  geom_errorbar( aes(x=reorder(frag1, meanboost), ymin=meanboost-sqrt(variance), ymax=meanboost+sqrt(variance), colour=CorrPvalue<0.05), width=0.4, alpha=0.9, size=1.3)+
  theme_bw()+
  labs(caption = "67 Klf2 locus cREs (combinations>5)")+
  coord_flip()
```

```{r BarplotVariancepvalue2, fig.height=15, fig.width=7}
all_comball%>%
  filter(class=="EP")%>%
  filter(strand2=="+")%>%
  mutate(frag1=paste0(frag1,"_", DAT))%>%
  mutate(frag2=paste0(frag2,"_", DAT))%>%
  mutate(id2=paste0(id2,"_", DAT))%>%
  filter(!BoostPerPmed%in%c(NA,NaN,Inf,-Inf))%>%
     filter(DAT=="klf2")%>%
  group_by(frag1, frag2)%>%
  summarise(BoostPerPmedAvgOrien=mean(BoostPerPmed))%>%
  filter(frag1%in%WelchTestEnhancers$frag1)%>%
 left_join(WelchTestEnhancers%>%mutate(CorrPvalue=p.adjust(pvalue, "BH")), by=c("frag1"))%>%
  left_join(VarPErE_AvgOr_onlyVar, by=c("frag1"))%>%
    mutate(frag1=sapply(str_split(frag1, "_"), function(x) x[1]))%>% 
  ggplot(aes(reorder(frag1, meanboost),BoostPerPmedAvgOrien, colour=CorrPvalue<0.05))+
  geom_hline(yintercept = 0, alpha=0.4)+
  geom_quasirandom(size=1.2, varwidth = TRUE, alpha=0.7)+
  stat_summary(fun.y = mean, width = 0.5, linetype = "solid", colour="black", geom = "crossbar")+
  #geom_bar(stat="identity")+
  #geom_crossbar( aes(x=reorder(frag1, meanboost),y=meanboost, ymin=meanboost-sqrt(variance), ymax=meanboost+sqrt(variance), colour=CorrPvalue<0.05), width=0.4, alpha=0.4, size=0.5)+
  #geom_crossbar( aes(x=reorder(frag1, meanboost),y=meanboost, ymin=meanboost, ymax=meanboost), colour="black", width=0.4, alpha=1, size=0.5)+
  #geom_point(aes(reorder(frag1, meanboost),meanboost, colour=CorrPvalue<0.05))+
  theme_bw(base_size = 15)+
  labs(caption = "67 Klf2 locus cREs (combinations>5)")+
  coord_flip()

WelchTestEnhancers%>%
  mutate(CorrPvalue=p.adjust(pvalue, "BH"))%>%
  mutate(Axis="enhancers")%>%
  left_join(VarPErE_AvgOr_onlyVar, by=c("frag1"))%>%
  mutate(DAT=sapply(str_split(frag1, "_"), function(x) x[2]))%>%
  mutate(frag1=sapply(str_split(frag1, "_"), function(x) x[1]))%>% 
  filter(DAT=="klf2")%>%
  ggplot(aes(reorder(frag1, meanboost),meanboost))+
  #geom_quasirandom(size=0.5, varwidth = TRUE, dodge.width=0.4,aes(color=CorrPvalue<0.05))+
  geom_bar(stat="identity")+
  geom_errorbar( aes(x=reorder(frag1, meanboost), ymin=meanboost-sqrt(variance), ymax=meanboost+sqrt(variance), colour=CorrPvalue<0.05), width=0.4, alpha=0.9, size=1.3)+
  theme_bw(base_size = 15)+
  labs(caption = "67 Klf2 locus cREs (combinations>5)")+
  coord_flip()
```


# Save data
```{r}
# save welch selectivity dataset


save('WelchTestEnhancers',
     file = '~/mydata/GitLab/EPCombinations/data/MMA20210811_SelectivityWelchEnh.RData')
```
```



